---
title: "20250304_xenium_ricover_revision2"
output: html_document
date: "2025-03-04"
---

```{r}
options(bitmapType="cairo")
```

#Set wd
```{r}
setwd("/g/huber/users/czernilo/proj/2021_LNS012/")
out_dir_obj <- "/g/huber/users/czernilo/proj/2021_LNS012/seurat_obj/"
out_dir_plots <- "/g/huber/users/czernilo/proj/2021_LNS012/output/"
out_dir_de <- "/g/huber/users/czernilo/proj/2021_LNS012/DE_analysis/"
out_dir_figs <- "/g/huber/users/czernilo/proj/2021_LNS012/output/fig/"
out_dir_xenium <- "/g/huber/users/czernilo/proj/2021_LNS012/xenium/"
```

#Annotation colors
```{r}
annotation_colors <- list("patient"= c("DLBCL-1"="#33691E",
                                        "DLBCL-2"="#55882f",
                                        "DLBCL-3"="#689F38",
                                        "DLBCL-4"="#7CB342",
                                        "DLBCL-5"="#8BC34A",
                                        "DLBCL-6"="#9CCC65",
                                        "DLBCL-7"="#AED581",
                                        "DLBCL-8"="#C5E1A5",
                                        "FL-1"="#86B0D3",
                                        "FL-2"="#5D9FC9",
                                        "FL-3"="#318DBE",
                                        "FL-4"="#0D75B2",
                                        "FL-5"="#04649C",
                                        "FL-6"="#03507E",
                                        "rLN-1"="#40004B",
                                        "rLN-2"="#71267D",
                                        "rLN-3"="#9263A3",
                                        "rLN-4"="#B696C5",
                                        "rLN-5"="#D9C2DE"),
                          "Type_binned"= c("rLN"= "#8D60B5",
                                            "FL"= "#8BCDE0",
                                            "DLBCL"= "#67B86E"),
                          "Type_binned_v2"=c("Healthy"= "#8D60B5",
                                             "FL"= "#8BCDE0",
                                             "DLBCL"= "#67B86E"),
                          "Type_binned_v3"=c("Tonsil"= "#8D60B5",
                                             "FL"= "#8BCDE0",
                                             "DLBCL"= "#67B86E"),
                          "Clusters_final"=c("artBEC"="#F31515",
                                                    "bLEC"="#C2A946",
                                                    "capBEC"="#F5818D",
                                                    "cLEC"="#FF9709",
                                                    "collectLEC"="#FFF16A",
                                                    "DLBCL-B"="#83CAE8",
                                                    "FDC"="#088B1A",
                                                    "FL-B"="#0EBCE8",
                                                    "fLEC"="gold1",
                                                    "Healthy B"="#0C80ED",
                                                    "MRC/IFR"="#A1D74F",
                                                    "LAF"="#2AC200",
                                                    "med-FRC"="#005C4C",
                                                    "msLEC"="#F5C56C",
                                                    "Myeloid"="snow3",
                                                    "PC"="#0330E3",
                                                    "PvC"="#659445",
                                                    "SMC"="#4E7329", #"#BEDEA6",
                                                    "T-FH"="mediumorchid1",
                                                    "T-H"="magenta4",
                                                    "T-PR"="purple4",
                                                    "T-REG"="darkorchid1",
                                                    "T-TOX"="mediumpurple1",
                                                    "TRC"="#96B577",
                                                    #"tipBEC"="hotpink1", 
                                                    "LA-BEC"="hotpink1",
                                                    "valveLEC"="#9E770E",
                                                    "vnBEC"="#9C194B"),
                          "Clusters"=c("artBEC"="tan1",
                                       "capBEC"="sienna3",
                                       "CCL19hi TRC"="forestgreen",
                                       "CCR7+ DC"="red4",
                                       "CD146- PvC"="#AAA64E",
                                       "CD34+ SC"="limegreen",
                                       "CD36+ PvC"="#6F7520",
                                       "CD8+ T"="darkorchid1",
                                       "cLEC"="gold1",
                                       "CXCL9+ TRC"="olivedrab3",
                                       "DLBCL-1 B"="cadetblue3",
                                       "DLBCL-1 SC"="olivedrab",
                                       "DLBCL-2 B"="deepskyblue1",
                                       "DLBCL-3 B"="skyblue",
                                       "DLBCL-5 B"="steelblue4",
                                       "DLBCL-6 B"="lightblue2",
                                       "DLBCL-7 B"="steelblue1",
                                       "DLBCL-8 B"="dodgerblue1",
                                       "ECM-hi SC"="chartreuse3",
                                       "FDC"="greenyellow",
                                       "FL B"="paleturquoise4",
                                       "FL-1 B"="turquoise3",
                                       "FL-2 B"="mediumaquamarine",
                                       "FL-2/3 SC"="palegreen4",
                                       "FL-3 B"="darkolivegreen3",
                                       "FL-4 B"="lightseagreen",
                                       "FL-4/6 B"="cyan4",
                                       "FL-5 B"="turquoise",
                                       "fLEC"="yellow3",
                                       "LYVE1+ LEC"="khaki",
                                       "Mø"="hotpink",
                                       "Naive B"="royalblue1",
                                       "NG2+ PvC"="#B5BF6B",
                                       "NK"="blueviolet",
                                       "NKT"="mediumpurple1",
                                       "PC"="blue3",
                                       "pDC"="tomato",
                                       "T-fh"="mediumorchid1",
                                       "T-h"="magenta4",
                                       "T-reg"="purple4",
                                       "vnBEC"="darkorange1"),
                          "binned_clusters"=c("artBEC"="#F31515",
                                              "capBEC"="#F5818D",
                                              "CD34+ SC"="#A9E075",
                                              "cLEC"="#FF9709",
                                              "DLBCL-B"="#83CAE8",
                                              "FDC"="#21D986",
                                              "FL-B"="#0EBCE8",
                                              "fLEC"="gold1",
                                              "Healthy B"="#0C80ED",
                                              "LAF"="#2AC200",
                                              "msLEC"="#F5C56C",
                                              "Myeloid"="#C7D6A1",
                                              "PC"="#0330E3",
                                              "PvC"="#71A73C",
                                              "T-FH"="mediumorchid1",
                                              "T-H"="magenta4",
                                              "T-PR"="purple4",
                                              "T-REG"="darkorchid1",
                                              "T-TOX"="mediumpurple1",
                                              "TRC"="#0B8B1A",
                                              "vnBEC"="#9C194B"),
                          "Clusters_Abe"=c("artBEC"="#F31515",
                                           "capBEC"="#F5818D",
                                           "advSC"="#A9E075",
                                           "cLEC"="#FF9709",
                                           "DLBCL-B"="#83CAE8",
                                           "FDC"="#21D986",
                                           "FL-B"="#0EBCE8",
                                           "fLEC"="gold1",
                                           "Healthy B"="#0C80ED",
                                           "LAF"="#2AC200",
                                           "msLEC"="#F5C56C",
                                           "Myeloid"="#C7D6A1",
                                           "valveLEC"="grey70",
                                           "AGT-SC"="#DBF227", 
                                           "MRC"="#005C53", 
                                           "PC"="#0330E3",
                                           "PvC/SMC"="#71A73C",
                                           "T-FH"="mediumorchid1",
                                           "T-H"="magenta4",
                                           "T-PR"="purple4",
                                           "T-REG"="darkorchid1",
                                           "T-TOX"="mediumpurple1",
                                           "TRC"="#0B8B1A",
                                           "vnBEC"="#9C194B", 
                                           "tipBEC"="grey70", 
                                           "bLEC"="grey70", 
                                           "collectLEC"="grey70"),
                          "binned_clusters2"=c("artBEC"="#F31515",
                                               "capBEC"="#F5818D",
                                               "CD34+ SC"="#A9E075",
                                               "cLEC"="#FF9709",
                                               "FDC"="#21D986",
                                               "fLEC"="gold1",
                                               "LAF"="#2AC200",
                                               "msLEC"="#F5C56C",
                                               "PvC"="#71A73C",
                                               "TRC"="#0B8B1A",
                                               "vnBEC"="#9C194B"),
                          "Clusters_paper"=c("BEC"="#F5818D",
                                                    "LEC"="#FF9709",
                                                    "DLBCL-B"="#83CAE8",
                                                    "FL-B"="#0EBCE8",
                                                    "Healthy B"="#0C80ED",
                                                    "Myeloid"="snow3",
                                                    "PC"="#0330E3",
                                                    "FRC"="#BEDEA6",
                                                    "FDC"="#088B1A",
                                                    "rFRC"="#2AC200",
                                                    "T-FH"="mediumorchid1",
                                                    "T-H"="magenta4",
                                                    "T-PR"="purple4",
                                                    "T-REG"="darkorchid1",
                                                    "T-TOX"="mediumpurple1",
                                                    "rBEC"="hotpink1"),
                          "gate"=c("DLBCL-B"="#83CAE8",
                                   "FL-B"="#0EBCE8",
                                   "fLEC"="gold1",
                                   "Healthy B"="#0C80ED",
                                   "Myeloid"="#C7D6A1",
                                   "PC"="#0330E3",
                                   "T-FH"="mediumorchid1",
                                   "T-H"="magenta4",
                                   "T-PR"="purple4",
                                   "T-REG"="darkorchid1",
                                   "T-TOX"="mediumpurple1",
                                   "DN"="#71A73C", 
                                   "MSC"="#2AC200", 
                                   "LEC"="#FF9709",
                                   "BEC"="#F5818D"),
                          "Clusters_gates"=c("artBEC"="#F31515",
                                             "capBEC"="#F5818D",
                                             "CD34+ SC"="#A9E075",
                                             "cLEC"="#FF9709",
                                             "DLBCL-B"="#83CAE8",
                                             "FDC"="#21D986",
                                             "FL-B"="#0EBCE8",
                                             "fLEC"="gold1",
                                             "Healthy B"="#0C80ED",
                                             "LAF"="#2AC200",
                                             "msLEC"="#F5C56C",
                                             "Myeloid"="#C7D6A1",
                                             "PC"="#0330E3",
                                             "PvC"="#71A73C",
                                             "T-FH"="mediumorchid1",
                                             "T-H"="magenta4",
                                             "T-PR"="purple4",
                                             "T-REG"="darkorchid1",
                                             "T-TOX"="mediumpurple1",
                                             "TRC"="#0B8B1A",
                                             "vnBEC"="#9C194B",
                                             "DN"="#71A73C", 
                                             "MSC"="#2AC200", 
                                             "LEC"="#FF9709",
                                             "BEC"="#9C194B"),
                          "FACS_Cluster"=c("BEC"="#E24275", 
                                           "DN"="#5F7C0B",
                                           "LEC"="#E67B09",
                                           "MSC"="#72C216"),
                          "Healthy_Abe"=c("msLEC"="#E65C17",
                                          "bLEC"="#FF862B",
                                          "cLEC"="#FFAA2B",
                                          "fLEC"="#F1BA2B",
                                          "valveLEC"="#FFDD1F"),
                          "features"=c(pals::parula(1000))
)
```

#Load libraries
```{r}
#library(aws.s3)

library(Seurat)
library(clustree)
library(ggplot2)
library(ggbeeswarm)
library(ggpubr)
library(gdata)

#library(pheatmap)
library(ComplexHeatmap)
#library(dittoSeq)
library(viridis)
library(RColorBrewer)
library(pals)
#library(wesanderson)
library(dplyr)
library(tibble)
library(stringr)
library(readxl)
library(patchwork)


library(org.Hs.eg.db)
library(GO.db)
library(biomaRt) #convert ms to human


library(limma) #strsplit2
#library(EnhancedVolcano)
library(clusterProfiler)
```

#Load cores
```{r}
genes <- read.csv("/g/huber/users/czernilo/proj/2021_LNS012/xenium/core1_genes.csv")

fl <- readRDS("/g/huber/users/czernilo/proj/2021_LNS012/xenium/core1_wFollicles_nuclear.rds")
head(rownames(fl@assays$RNA))
rownames(fl@assays$RNA) <- genes$X
head(rownames(fl@assays$RNA))
fl$type <- "FL"

dlbcl2 <- readRDS("/g/huber/users/czernilo/proj/2021_LNS012/xenium/core3_nuclear.rds")
head(rownames(dlbcl2@assays$RNA))
rownames(dlbcl2@assays$RNA) <- genes$X
head(rownames(dlbcl2@assays$RNA))
dlbcl2$type <- "DLBCL"
```

#Merge
```{r}
DefaultAssay(fl) <- "RNA"
DefaultAssay(dlbcl2) <- "RNA"

# Manually merge count matrices before merging Seurat objects
merged_counts <- cbind(GetAssayData(fl, layer = "counts"), GetAssayData(dlbcl2, layer = "counts"))
merged_meta <- rbind(fl@meta.data, dlbcl2@meta.data)

# Create a new Seurat object with the merged count matrix
merged_obj <- CreateSeuratObject(counts = merged_counts, meta.data = merged_meta, assay = "RNA")
dim(merged_obj@assays$RNA)
```


#Filter
```{r}
VlnPlot(merged_obj, features = "nCount_RNA", pt.size = 0.1) + NoLegend() + geom_hline(yintercept = 50)

merged_obj <- subset(merged_obj, subset = nCount_RNA > 50)
```
#Log-normalize and SCT
```{r}
DefaultAssay(merged_obj) <- "RNA"
merged_obj <- NormalizeData(merged_obj, assay = "RNA")
merged_obj <- SCTransform(merged_obj, verbose = FALSE)

```

#Clustering
```{r}
merged_obj <- RunPCA(merged_obj, npcs = 30, features = rownames(merged_obj))
merged_obj <- RunUMAP(merged_obj, dims = 1:30)
merged_obj <- FindNeighbors(merged_obj, reduction = "pca", dims = 1:30)
merged_obj <- FindClusters(merged_obj, resolution = 0.3)
```

```{r}
DimPlot(merged_obj, group.by = "seurat_clusters", label = T)
DimPlot(merged_obj, group.by = "type")
```

#Annotate cell types
```{r, fig.height=20, fig.width=20}
markers <- FindAllMarkers(merged_obj, only.pos = TRUE, min.pct = 0.2, assay = "RNA") 
write.csv(markers, "/g/huber/users/czernilo/proj/2021_LNS012/xenium/markers.csv", row.names = FALSE)
top15 <- markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_log2FC)

pdf(paste0(out_dir_xenium, "xenium_markers_heatmap.pdf"), width = 15, height = 15)
DoHeatmap(merged_obj, features = c(unique(top10$gene)), draw.line=T, label=T) + 
       theme(axis.text.y=element_text(size = 15)) #+ scale_fill_gradient(colors = c("blue", "white", "red")))
dev.off()
```

```{r}
merged_obj@meta.data <- mutate(merged_obj@meta.data, cell_type = case_when(seurat_clusters %in% c(0)~ "FL-B",
                                                         seurat_clusters %in% c(1)~ "DLBCL-B",
                                                         seurat_clusters %in% c(2)~ "T",
                                                         seurat_clusters %in% c(3)~ "FL-B prol",
                                                         seurat_clusters %in% c(4)~ "DLBCL-B prol",
                                                         seurat_clusters %in% c(5)~ "Myeloid",
                                                         seurat_clusters %in% c(6)~ "FRC",
                                                         seurat_clusters %in% c(7)~ "rFRC",
                                                         seurat_clusters %in% c(8)~ "BEC",
                                                         seurat_clusters %in% c(9)~ "FDC",
                                                         seurat_clusters %in% c(10)~ "Myeloid",
                                                         seurat_clusters %in% c(11)~ "weird"))

DimPlot(merged_obj, group.by = "cell_type", label = T) + ggtitle("merged")

```

##Subcluster unclear cells
```{r}
bec <- subset(merged_obj, cell_type %in% c("BEC", "weird")) 
DefaultAssay(bec) <- "SCT"
```

```{r}
bec <- RunPCA(bec, npcs = 30, features = rownames(merged_obj))
bec <- RunUMAP(bec, dims = 1:30)
bec <- FindNeighbors(bec, reduction = "pca", dims = 1:30)
bec <- FindClusters(bec, resolution = 0.2)
```

```{r, fig.height=15, fig.width=15}
markers <- FindAllMarkers(bec, only.pos = TRUE, min.pct = 0.2, assay = "RNA") 

top15 <- markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_log2FC)

DoHeatmap(bec, features = c(unique(top15$gene)), draw.line=T, label=T) + 
       theme(axis.text.y=element_text(size = 15)) #+ scale_fill_gradient(colors = c("blue", "white", "red")))
```

```{r}
bec@meta.data <- mutate(bec@meta.data, cell_type2 = case_when(seurat_clusters %in% c(0)~ "BEC",
                                                         seurat_clusters %in% c(1)~ "FRC",
                                                         seurat_clusters %in% c(2)~ "FL-B",
                                                         seurat_clusters %in% c(3)~ "FRC"))

DimPlot(bec, group.by = "cell_type2")
```

##Rename unclear clusters in original obj
```{r}
#Extract cell names that belong to the clusters of interest
bec_new <- rownames(bec@meta.data[which(bec@meta.data$seurat_clusters %in% c(0)), ])
b_new <- rownames(bec@meta.data[which(bec@meta.data$seurat_clusters %in% c(2)), ])
frc_new <- rownames(bec@meta.data[which(bec@meta.data$seurat_clusters %in% c(1,3)), ])

#Define their new name
bec_new_name <- rep("BEC", length(bec_new))
names(bec_new_name) <- bec_new

b_new_name <- rep("FL-B", length(b_new))
names(b_new_name) <- b_new

frc_new_name <- rep("FRC", length(frc_new))
names(frc_new_name) <- frc_new

#If you have more than 1 subclusters renamed, you need to merge 
rename_subclusters <-  c(bec_new_name, b_new_name, frc_new_name)

#Add metadata information to object
merged_obj <- AddMetaData(merged_obj, 
                        metadata = rename_subclusters, 
                        col.name = "subcluster")
merged_obj@meta.data$cell_type2 <- ifelse(!is.na(merged_obj@meta.data$subcluster), 
                                        merged_obj@meta.data$subcluster, 
                                        merged_obj@meta.data$cell_type)

merged_obj@meta.data$cell_type <- merged_obj@meta.data$cell_type2
merged_obj@meta.data$cell_type2 <- NULL
```

```{r}
DimPlot(merged_obj, group.by = "cell_type", label=T) + ggtitle("merged")

```
##Subcluster T cells
```{r}
t <- subset(merged_obj, cell_type %in% c("T")) 
DefaultAssay(t) <- "SCT"
```
```{r}
t <- RunPCA(t, npcs = 30, features = rownames(merged_obj))
t <- RunUMAP(t, dims = 1:30)
t <- FindNeighbors(t, reduction = "pca", dims = 1:30)
t <- FindClusters(t, resolution = 0.2)
```

```{r, fig.height=15, fig.width=15}
markers <- FindAllMarkers(t, only.pos = TRUE, min.pct = 0.2, assay = "RNA") 
top15 <- markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_log2FC)

DoHeatmap(t, features = c(unique(top15$gene)), draw.line=T, label=T) + 
       theme(axis.text.y=element_text(size = 15)) #+ scale_fill_gradient(colors = c("blue", "white", "red")))
```

```{r}
t@meta.data <- mutate(t@meta.data, cell_type2 = case_when(seurat_clusters %in% c(0)~ "T-FH",
                                                         seurat_clusters %in% c(1)~ "T-REG",
                                                         seurat_clusters %in% c(2)~ "T-TOX",
                                                         seurat_clusters %in% c(3)~ "T-H"))

DimPlot(t, group.by = "cell_type2")
FeaturePlot(t, features = "CD8A")
FeaturePlot(t, features = "CD4")
FeaturePlot(t, features = "CXCR5")

```

##Rename T clusters in original object
```{r}
#Extract cell names that belong to the clusters of interest
tox_new <- rownames(t@meta.data[which(t@meta.data$seurat_clusters %in% c(2)), ])
reg_new <- rownames(t@meta.data[which(t@meta.data$seurat_clusters %in% c(1)), ])
fh_new <- rownames(t@meta.data[which(t@meta.data$seurat_clusters %in% c(0)), ])
h_new <- rownames(t@meta.data[which(t@meta.data$seurat_clusters %in% c(3)), ])

#Define their new name
tox_new_name <- rep("T-TOX", length(tox_new))
names(tox_new_name) <- tox_new

reg_new_name <- rep("T-REG", length(reg_new))
names(reg_new_name) <- reg_new

fh_new_name <- rep("T-FH", length(fh_new))
names(fh_new_name) <- fh_new

h_new_name <- rep("T-H", length(h_new))
names(h_new_name) <- h_new

#If you have more than 1 subclusters renamed, you need to merge 
rename_subclusters <-  c(tox_new_name, reg_new_name, fh_new_name, h_new_name)

#Add metadata information to object
merged_obj <- AddMetaData(merged_obj, 
                        metadata = rename_subclusters, 
                        col.name = "subcluster")
merged_obj@meta.data$cell_type2 <- ifelse(!is.na(merged_obj@meta.data$subcluster), 
                                        merged_obj@meta.data$subcluster, 
                                        merged_obj@meta.data$cell_type)

merged_obj@meta.data$cell_type <- merged_obj@meta.data$cell_type2
merged_obj@meta.data$cell_type2 <- NULL
```

##Tidy annotation
```{r}
merged_obj@meta.data <- mutate(merged_obj@meta.data, cell_type_binned = case_when(cell_type %in% c("FL-B prol", "FL-B")~ "FL-B",
                                                         cell_type %in% c("DLBCL-B prol", "DLBCL-B")~ "DLBCL-B",
                                                         cell_type %in% c("T-TOX")~ "T-TOX",
                                                         cell_type %in% c("T-REG")~ "T-REG",
                                                         cell_type %in% c("T-H")~ "T-H",
                                                         cell_type %in% c("T-FH")~ "T-FH",
                                                         cell_type %in% c("Myeloid")~ "Myeloid",
                                                         cell_type %in% c("FRC")~ "FRC",
                                                         cell_type %in% c("rFRC")~ "rFRC",
                                                         cell_type %in% c("BEC")~ "BEC",
                                                         cell_type %in% c("FDC")~ "FDC"))

```

#Save
```{r}
saveRDS(merged_obj, "/g/huber/users/czernilo/proj/2021_LNS012/xenium/xenium_merged_seurat.rds")
```

#Figure DimPlot
```{r Panel B left, fig.height=8, fig.width=12}
data <- Embeddings(object = merged_obj[["umap"]])[(colnames(merged_obj)), c(1, 2)]
data <- as.data.frame(data)
data$cell_type_binned <- merged_obj$cell_type_binned
data$Type_binned <- merged_obj$type
data$cell_type_binned <- merged_obj$cell_type_binned

col_vec_umap <- annotation_colors$Clusters_paper

col_vec <-  unlist(lapply(data$cell_type_binned, function(x){index=which(x==names(col_vec_umap))
col_vec_umap[index]}))
data$color_patient <- col_vec

levels = c("FL-B", "DLBCL-B", "T-H", "T-FH", "T-REG", "T-TOX", "Myeloid", "BEC", "FDC", "FRC", "rFRC")
data$cell_type_binned <- factor(x = data$cell_type_binned, levels = levels) #change order of factor levels for better visualization
```


```{r Panel B left, fig.height=8, fig.width=12, fig.cap="UMAP of RNA sequencing data"}
pdf(paste0(out_dir_figs, "xenium_umap.pdf"), width =7, height = 7)
ggplot(data,aes(umap_1, umap_2)) + 
  ggforce::geom_voronoi_tile(aes(fill = cell_type_binned, group = -1L), max.radius =0.05) +
  theme_void()+
  theme(legend.position = "bottom")+ 
  scale_fill_manual(values= col_vec) 
dev.off()
```

#Subcluster FDC/FRC/rFRC (not shown, exclude code)
```{r}
frc <- subset(merged_obj, cell_type %in% c("FRC", "rFRC", "FDC")) 
DefaultAssay(frc) <- "SCT"
```
```{r}
frc <- RunPCA(frc, npcs = 30, features = rownames(merged_obj))
frc <- RunUMAP(frc, dims = 1:30)
frc <- FindNeighbors(frc, reduction = "pca", dims = 1:30)
frc <- FindClusters(frc, resolution = 0.2)
```

##UMAP FRC paper
```{r}
levels <- c("FDC", "FRC", "rFRC")
cols <- c("#088B1A", "#BEDEA6", "#2AC200")

dim <- DimPlot(frc, group.by = "cell_type_binned", label = F, label.box = F, repel = F, cols = cols, raster = F, pt.size = 1) + theme(legend.position="right")
dim$data$cell_type_binned <- factor(x = dim$data$cell_type_binned, levels = levels) #change order of factor levels for better visualization

pdf(paste0(out_dir_figs, "xenium_frc_umap.pdf"), width = 7, height = 5)
dim
dev.off()
```

```{r}
frc@meta.data$type <- factor(frc@meta.data$type, levels = c("FL", "DLBCL"))

pdf(paste0(out_dir_figs, "xenium_frc_homeo_vln.pdf"), width = 4, height = 4)
VlnPlot(frc, features = "homeo_score2", group.by = "type", pt.size = 0, y.max = 5.5) +
      #stat_compare_means(comparisons = comparisons, method = "wilcox.test", label = "p.signif", ref.group = ".all.", hide.ns = FALSE)
      geom_boxplot(inherit.aes = T, width=0.1, outlier.size = 0, alpha = 0.6, notch = FALSE) + 
      geom_hline(yintercept = mean(frc$homeo_score2), linetype = 2) + # Add horizontal line at base mean
      #scale_y_continuous(limits=c(-1, 3.5), breaks=c(-1,0,1,2,3))+
      labs(title = "Homeostatic chemokines", x = "", y = "Module Score\n") +
      theme(plot.title = element_text(size = 15, color = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_rect(color = "black"), 
            aspect.ratio=1.5,
            axis.text.y = element_text(size = 12, color = "black"),
            axis.title.x = element_blank(),
            axis.title.y = element_text(size = 15, color = "black"), 
            axis.text.x = element_text(size = 15, color = "black", angle = 0, hjust = 0.5)) + 
      #scale_x_discrete(labels= c("NHL-associated", "healthy-associated")) +
    scale_fill_manual(values=c("#8BCDE0", "#67B86E")) +
      NoLegend()
dev.off()
```

#Marker and chemokine expression across cell types
##Violin
```{r}
chemokines <- c("CXCL13", "CXCL12", "CCL21", "CCL19", "CXCL9", "CXCL10", "CXCL11")

levels = c("FL-B", "DLBCL-B", "T", "Myeloid", "BEC", "FRC", "FDC", "rFRC")
merged_obj@meta.data$cell_type_binned <- factor(x = merged_obj@meta.data$cell_type_binned, levels = levels)

VlnPlot(merged_obj, features= chemokines, group.by = "cell_type_binned", fill.by = "ident", pt.size = 0, stack = T, flip = T) + NoLegend() + xlab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12, face = "plain", margin = margin(5, 0, 0, 0)))

```
##Heatmaps
```{r}
markers <- c("CXCL13", "CXCL12", "CCL21", "CXCL9", "CXCL10", "CXCL11")
Idents(merged_obj) <- "cell_type_binned"
avg_expr <- AverageExpression(merged_obj, assays = "RNA", features = markers, return.seurat = F) 
avg_expr <- as.data.frame(avg_expr[["RNA"]]) 

#Reorder
avg_expr <- avg_expr[,  c("FDC", "FRC", "rFRC", "BEC", "Myeloid", "T-TOX", "T-REG", "T-FH", "T-H", "FL-B", "DLBCL-B")]
avg_expr <- avg_expr[c("CXCL13", "CXCL12", "CCL21", "CXCL9", "CXCL10", "CXCL11"),]

#Change rownames
#rownames(mat_scores_reclust) <- c("ABEC", "caBEC", "aBEC", "tBEC", "cBEC", "caHEV", "aHEV", "hHEV", "CXCL10-HEV", "VBEC")

#Scale
scaled_avg_expr <- scale(t(avg_expr))
#scaled_avg_expr <- t(scaled_avg_expr)
```
```{r, fig.height=10, fig.width=10}
#pdf(paste0(out_dir_figs, "xenium_chemokine_heatmap.pdf"), width = 5, height = 5)
h2 <- Heatmap(scaled_avg_expr, 
        cluster_columns = F, 
        cluster_rows = F,
        #col = rev(brewer.pal(n = 8, name = 'RdBu')),
        col = circlize::colorRamp2(c(-2, 0, 2), c("#007EA7", "white", "#AD2E24")),
        heatmap_legend_param = list(
          title = "Scaled expression", 
          color_bar = "continuous",
          at = c(-2, 0, 2),
          legend_direction = "horizontal", 
          title_gp = gpar(fontsize = 10, fontface = "bold")),
        width = ncol(scaled_avg_expr)*unit(5, "mm"), 
        height = nrow(scaled_avg_expr)*unit(5, "mm"), 
        row_names_gp = grid::gpar(fontsize = 10),
        column_names_gp = grid::gpar(fontsize = 10),
        rect_gp = gpar(col = "black", lwd = 1),
        column_dend_height = unit(10, "mm"),
        row_dend_width = unit(10, "mm")
        )
#dev.off()
```

```{r}
markers <- c("CR2", "PDGFRB", "THY1", "PECAM1", "ITGAX", "CD14", "CD163", "CD8A", "FOXP3", "ICOS", "CXCR5", "TCF7", "TRBC2", "CD19")
Idents(merged_obj) <- "cell_type_binned"
avg_expr <- AverageExpression(merged_obj, assays = "RNA", features = markers, return.seurat = F) 
avg_expr <- as.data.frame(avg_expr[["RNA"]]) 

#Reorder
avg_expr <- avg_expr[,  c("FDC", "FRC", "rFRC", "BEC", "Myeloid", "T-TOX", "T-REG", "T-FH", "T-H", "FL-B", "DLBCL-B")]
avg_expr <- avg_expr[c("CR2", "PDGFRB", "THY1", "PECAM1", "ITGAX", "CD14", "CD163", "CD8A", "FOXP3", "ICOS", "CXCR5", "TCF7", "TRBC2", "CD19"),]

#Scale
scaled_avg_expr <- scale(t(avg_expr))
```

```{r, fig.height=10, fig.width=10}
h1 <- Heatmap(t(scaled_avg_expr), 
        cluster_columns = F, 
        cluster_rows = F,
        #col = rev(brewer.pal(n = 8, name = 'RdBu')),
        col = circlize::colorRamp2(c(-2, 0, 2), c("#007EA7", "white", "#AD2E24")),
        heatmap_legend_param = list(
          title = "Scaled expression", 
          color_bar = "continuous",
          at = c(-2, 0, 2),
          legend_direction = "horizontal", 
          title_gp = gpar(fontsize = 10, fontface = "bold")),
        width = ncol(scaled_avg_expr)*unit(5, "mm"), 
        height = nrow(scaled_avg_expr)*unit(5, "mm"), 
        row_names_gp = grid::gpar(fontsize = 10),
        column_names_gp = grid::gpar(fontsize = 10),
        rect_gp = gpar(col = "black", lwd = 1),
        column_dend_height = unit(10, "mm"),
        row_dend_width = unit(10, "mm")
        )
```

```{r}
pdf(paste0(out_dir_figs, "xenium_chemokine_marker_heatmap.pdf"), width = 8, height = 8)
h1 + h2
dev.off()
```

#Stacked bar plot FRCs
```{r}
# Subset metadata for relevant cell types
meta_df <- as.data.frame(merged_obj@meta.data) %>%
  filter(cell_type_binned %in% c("FDC", "FRC", "rFRC"))

# Count total number of cells per cell type and disease type
cell_counts <- meta_df %>%
  group_by(cell_type_binned, type) %>%
  summarise(count = n(), .groups = "drop")  # Absolute cell count

# Compute proportions for relative comparison
cell_counts <- cell_counts %>%
  group_by(type) %>%
  mutate(proportion = count / sum(count))  # Normalize within each disease type

# Convert factors for ordered plotting
cell_counts$cell_type_binned <- factor(cell_counts$cell_type_binned, levels = c("FDC", "FRC", "rFRC"))
cell_counts$type <- factor(cell_counts$type, levels = c("FL", "DLBCL"))
```

```{r}
ggplot(cell_counts, aes(x = type, y = count, fill = cell_type_binned)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.6) +  # Side-by-side bars
  scale_fill_manual(values = c("#1f78b4", "#33a02c", "#e31a1c")) +  # Custom colors
  theme_bw() +
  labs(title = "Absolute Cell Numbers per Type",
       x = "Disease Type",
       y = "Cell Count",
       fill = "Cell Type") +
  theme(plot.title = element_text(size = 15, color = "black"),
        axis.text.x = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 14, color = "black"))
```

```{r}
p1 <- ggplot(cell_counts, aes(x = type, y = proportion, fill = cell_type_binned)) +
  geom_bar(stat = "identity", position = "fill", width = 0.5) +  # Stacked proportions
  scale_fill_manual(values = c("#088B1A", "#BEDEA6", "#2AC200")) + 
  theme_bw() +
  theme(plot.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "black"), 
        legend.position = "none",
        aspect.ratio = 1.5,
        axis.text.y = element_text(size = 12, color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank())
```

#Percent positive cells (not shown, exclude code)
##CXCL13
```{r}
# Define CXCL13+ status across all cell types
RidgePlot(merged_obj, features="CXCL13")
cxcl13_positive <- GetAssayData(merged_obj, assay = "RNA", slot = "data")["CXCL13", ] > 2
merged_obj$CXCL13_status <- ifelse(cxcl13_positive, "CXCL13+", "CXCL13-")

# Convert metadata to dataframe
meta_df <- as.data.frame(merged_obj@meta.data)

# Count total cells per `cell_type_binned`, `type`, and CXCL13 status
cxcl13_summary <- meta_df %>%
  group_by(cell_type_binned, type, CXCL13_status) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cell_type_binned, type) %>%
  mutate(proportion = count / sum(count))  # Compute proportion of CXCL13+ cells per type

# Order factors for consistency
cxcl13_summary$type <- factor(cxcl13_summary$type, levels = c("FL", "DLBCL"))
cxcl13_summary$cell_type_binned <- factor(cxcl13_summary$cell_type_binned, 
                                          levels = c("FDC", "FRC", "rFRC", "BEC", "Myeloid", "T-TOX", "T-REG", "T-FH", "T-H", "FL-B", "DLBCL-B"))
```

```{r}
ggplot(cxcl13_summary, aes(x = type, y = proportion, fill = CXCL13_status)) +
  geom_bar(stat = "identity", position = "fill", width = 0.5) +  # Stacked proportions
  #scale_fill_manual(values = c("#088B1A", "#BEDEA6", "#2AC200")) + 
  theme_bw() +
  theme(plot.title = element_text(size = 15, color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "black"), 
        aspect.ratio = 1.5,
        axis.text.y = element_text(size = 12, color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1))
```

```{r}
ggplot(cxcl13_summary, aes(x = type, y = proportion, fill = CXCL13_status)) +
  geom_bar(stat = "identity", position = "fill", width = 0.6) +  # Stacked proportions
  scale_fill_manual(values = c("#f9f9f9", "#AD2E24")) +  # Custom colors
  theme_bw() +
  facet_wrap(~ cell_type_binned, scales = "free_y") +  # Facet by cell type
  theme(plot.title = element_text(size = 15, color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "black"), 
        aspect.ratio = 1.5,
        axis.text.y = element_text(size = 12, color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1)) +
  labs(fill = "CXCL13 Status", title = "CXCL13+ Proportion Across Cell Types")
```

##CXCL9
```{r}
# Define CXCL9+ status across all cell types
RidgePlot(merged_obj, features="CXCL9")
cxcl9_positive <- GetAssayData(merged_obj, assay = "RNA", slot = "data")["CXCL9", ] > 2
merged_obj$CXCL9_status <- ifelse(cxcl9_positive, "CXCL9+", "CXCL9-")

# Convert metadata to dataframe
meta_df <- as.data.frame(merged_obj@meta.data)

# Count total cells per `cell_type_binned`, `type`, and CXCL9 status
cxcl9_summary <- meta_df %>%
  group_by(cell_type_binned, type, CXCL9_status) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cell_type_binned, type) %>%
  mutate(proportion = count / sum(count))  # Compute proportion of CXCL9+ cells per type

# Order factors for consistency
cxcl9_summary$type <- factor(cxcl9_summary$type, levels = c("FL", "DLBCL"))
cxcl9_summary$cell_type_binned <- factor(cxcl9_summary$cell_type_binned, 
                                          levels = c("FDC", "FRC", "rFRC", "BEC", "Myeloid", "T-TOX", "T-REG", "T-FH", "T-H", "FL-B", "DLBCL-B"))
```

```{r}
ggplot(cxcl9_summary, aes(x = type, y = proportion, fill = CXCL9_status)) +
  geom_bar(stat = "identity", position = "fill", width = 0.6) +  # Stacked proportions
  scale_fill_manual(values = c("#f9f9f9", "#AD2E24")) +  # Custom colors
  theme_bw() +
  facet_wrap(~ cell_type_binned, scales = "free_y") +  # Facet by cell type
  theme(plot.title = element_text(size = 15, color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "black"), 
        aspect.ratio = 1.5,
        axis.text.y = element_text(size = 12, color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1)) +
  labs(fill = "CXCL9 Status")
```

#Chemokine expression
##CXCL13
```{r}
chemokines <- c("CXCL13", "CXCL12", "CCL21", "CCL19", "CXCL9", "CXCL10", "CXCL11")

frc@meta.data$cell_type_binned <- factor(x = frc@meta.data$cell_type_binned, levels = c("FDC", "FRC", "rFRC"))
frc@meta.data$type <- factor(x = frc@meta.data$type, levels = c("FL", "DLBCL"))
comparisons <- list(c("FL", "DLBCL"))

VlnPlot(frc, features= chemokines, group.by = "cell_type_binned", fill.by = "ident", pt.size = 0, stack = T, flip = T) + NoLegend() + xlab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12, face = "plain", margin = margin(5, 0, 0, 0)))

p2 <- VlnPlot(frc, features = "CXCL13", group.by = "type", fill.by = "cell_type_binned", pt.size = 0, y.max = 8) +
      stat_compare_means(comparisons = comparisons, method = "wilcox.test", label = "p.signif", ref.group = ".all.", hide.ns = FALSE) +
      geom_boxplot(inherit.aes = T, width=0.1, outlier.size = 0, alpha = 0.6, notch = FALSE) + 
      #geom_hline(yintercept = mean(frc@assays$RNACXCL13), linetype = 2) + # Add horizontal line at base mean
      ggtitle("CXCL13") + 
      #scale_y_continuous(limits=c(-0.4, 1.2), breaks=c(-0.3,0,0.3,0.6,0.9))+
      theme(plot.title = element_blank(),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_rect(color = "black"), 
            aspect.ratio=1,
            axis.text.y = element_text(size = 12, color = "black"),
            axis.title.x = element_blank(),
            #axis.title.y = element_text(size = 15, color = "black", vjust = 1), 
            axis.title.y = element_blank(), 
            axis.text.x = element_text(size = 12, color = "black", angle = 0, hjust = 0.5)) + 
      #scale_x_discrete(labels= c("NHL-associated", "healthy-associated")) +
      scale_fill_manual(values = c("#8BCDE0", "#67B86E")) +
      NoLegend()
```

```{r}
pdf(paste0(out_dir_figs, "xenium_frc_cxcl13_bar.pdf"), width = 2, height = 4)
cowplot::plot_grid(p1, p2, nrow = 2, ncol = 1)
dev.off()
```

##Scored
```{r}
ifn_score <- c("CXCL9", "CXCL10", "CXCL11")
homeo_score <- c("CXCL13", "CXCL12", "CCL21")
comparisons <- list(c("FL", "DLBCL"))

scores <- list()
scores[[1]] <- ifn_score
scores[[2]] <- homeo_score

DefaultAssay(frc) <- "RNA"
frc <- AddModuleScore(object = frc, features = scores, ctrl = 5, name = c("ifn_score", "homeo_score"))
frc[['Scores']] <- CreateAssayObject(data = t(x = FetchData(object = frc, vars = c("ifn_score1", "homeo_score2"))))
frc@assays$Scores
```

```{r}
subset(frc, cell_type_binned %in% c("FRC")) %>% 
VlnPlot(features = "ifn_score1", group.by = "type", pt.size = 0, y.max = 2.2) +
      stat_compare_means(comparisons = comparisons, method = "wilcox.test", label = "p.signif", ref.group = ".all.", hide.ns = FALSE) +
      geom_boxplot(inherit.aes = T, width=0.1, outlier.size = 0, alpha = 0.6, notch = FALSE) + 
      geom_hline(yintercept = mean(frc$ifn_score1), linetype = 2) + # Add horizontal line at base mean
      ggtitle("Homeo") + 
      #scale_y_continuous(limits=c(-0.4, 1.2), breaks=c(-0.3,0,0.3,0.6,0.9))+
      theme(plot.title = element_text(size = 15, color = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_rect(color = "black"), 
            aspect.ratio=1,
            axis.text.y = element_text(size = 12, color = "black"),
            axis.title.x = element_blank(),
            #axis.title.y = element_text(size = 15, color = "black", vjust = 1), 
            axis.title.y = element_blank(), 
            axis.text.x = element_text(size = 15, color = "black", angle = 0, hjust = 0.5)) + 
      labs(x = "", y = "Module Score\n") +
      #scale_x_discrete(labels= c("NHL-associated", "healthy-associated")) +
      scale_fill_manual(values = c("#8BCDE0", "#67B86E")) +
      NoLegend()

VlnPlot(frc, features = "homeo_score2", group.by = "type", pt.size = 0, y.max = 2.2) +
      stat_compare_means(comparisons = comparisons, method = "wilcox.test", label = "p.signif", ref.group = ".all.", hide.ns = FALSE) +
      geom_boxplot(inherit.aes = T, width=0.1, outlier.size = 0, alpha = 0.6, notch = FALSE) + 
      geom_hline(yintercept = mean(comparisons$homeo_score2), linetype = 2) + # Add horizontal line at base mean
      ggtitle("Homeo") + 
      #scale_y_continuous(limits=c(-0.4, 1.2), breaks=c(-0.3,0,0.3,0.6,0.9))+
      theme(plot.title = element_text(size = 15, color = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_rect(color = "black"), 
            aspect.ratio=1,
            axis.text.y = element_text(size = 12, color = "black"),
            axis.title.x = element_blank(),
            #axis.title.y = element_text(size = 15, color = "black", vjust = 1), 
            axis.title.y = element_blank(), 
            axis.text.x = element_text(size = 15, color = "black", angle = 0, hjust = 0.5)) + 
      labs(x = "", y = "Module Score\n") +
      #scale_x_discrete(labels= c("NHL-associated", "healthy-associated")) +
      scale_fill_manual(values = c("#8BCDE0", "#67B86E")) +
      NoLegend()
```

#Visualize chemokines in space
##FL
```{r}
fl <- subset(merged_obj, type %in% c("FL"))
genes <- c("CXCL13", "CXCL12", "CCL21", "CCL19", "CXCL9", "CXCL10", "CXCL11", "CCR7", "CR2", "CR1", "PDGFRB", "THY1", "CXCR3")
metadata_fl <- as.data.frame(fl@meta.data)
DefaultAssay(fl) <- "RNA"
fl_df <-  cbind(metadata_fl, FetchData(fl, genes))
unique(fl_df$cell_type_binned)
```

###Cell types
```{r}
pdf(paste0(out_dir_figs, "xenium_fl_celltypes_full_spatial.pdf"), width = 4, height = 4)
ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned)) +
  # Plot the background points with #f0f0f0
  geom_point(data = subset(fl_df, cell_type_binned %in% c("BEC", "DLBCL-B", "Myeloid", "T-TOX", "T-REG", "T-FH", "T-H")),
             aes(alpha = 1), size = 0.25, color = "#f0f0f0") +
  
    # Plot the "FL-B" cells with alpha = 0
  geom_point(data = subset(fl_df, cell_type_binned == "FL-B"),
             aes(alpha = 0), size = 0.25, color = "#0EBCE8") +
  
  # Plot all other colored points with alpha = 1
  geom_point(data = subset(fl_df, cell_type_binned %in% c("FRC", "rFRC", "FDC")),
             aes(alpha = 1), size = 0.25) +
  
  scale_color_manual(values = c(
    "BEC" = "#f0f0f0",
    "Myeloid" = "#f0f0f0",
    "T-TOX" = "#f0f0f0",
    "T-REG"= "#f0f0f0", 
    "T-FH"= "#f0f0f0", 
    "T-H" = "#f0f0f0",
    "DLBCL-B" = "#f0f0f0",
    "FL-B" = "#0EBCE8",
    "FRC" = "#BEDEA6",
    "rFRC" = "#2AC200",
    "FDC" = "#088B1A"))+
  theme_bw() +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"))
dev.off()
```

```{r}
pdf(paste0(out_dir_figs, "xenium_fl_celltypes_zoom_spatial.pdf"), width = 4, height = 4)
ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned)) +
  geom_point(data = subset(fl_df, cell_type_binned %in% c("BEC", "DLBCL-B", "Myeloid", "T-TOX", "T-REG", "T-FH", "T-H")),
             aes(alpha = 1), size = 0.5, color = "#f0f0f0") +
  
    # Plot the "FL-B" cells with alpha = 0
  geom_point(data = subset(fl_df, cell_type_binned == "FL-B"),
             aes(alpha = 0.5), size = 0.5) +
  
  # Plot all other colored points with alpha = 1
  geom_point(data = subset(fl_df, cell_type_binned %in% c("FRC", "rFRC", "FDC")),
             aes(alpha = 1), size = 0.5) +
  
  theme_bw() +
  scale_color_manual(values = c(
    "BEC" = "#f0f0f0",
    "Myeloid" = "#f0f0f0",
    "T-TOX" = "#f0f0f0",
    "T-REG"= "#f0f0f0", 
    "T-FH"= "#f0f0f0", 
    "T-H" = "#f0f0f0",
    "DLBCL-B" = "#f0f0f0",
    "FL-B" = "#0EBCE8",
    "FRC" = "#BEDEA6",
    "rFRC" = "#2AC200",
    "FDC" = "#088B1A"
  )) +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)+
  xlim(2600, 3100) +
  ylim(4100, 4600) 
dev.off()
```
```{r}
pdf(paste0(out_dir_figs, "xenium_fl_celltypes_zoom_spatial_legend.pdf"), width = 4, height = 4)
ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned)) +
  geom_point(data = subset(fl_df, cell_type_binned %in% c("BEC", "DLBCL-B", "Myeloid", "T-TOX", "T-REG", "T-FH", "T-H")),
             aes(alpha = 1), size = 1, color = "#f0f0f0") +
  
    # Plot the "FL-B" cells with alpha = 0
  geom_point(data = subset(fl_df, cell_type_binned == "FL-B"),
             aes(alpha = 0.8), size = 1) +
  
  # Plot all other colored points with alpha = 1
  geom_point(data = subset(fl_df, cell_type_binned %in% c("FRC", "rFRC", "FDC")),
             aes(alpha = 1), size = 1) +
  
  theme_bw() +
  scale_color_manual(values = c(
    "BEC" = "#f0f0f0",
    "Myeloid" = "#f0f0f0",
    "T-TOX" = "#f0f0f0",
    "T-REG"= "#f0f0f0", 
    "T-FH"= "#f0f0f0", 
    "T-H" = "#f0f0f0",
    "DLBCL-B" = "#f0f0f0",
    "FL-B" = "#0EBCE8",
    "FRC" = "#BEDEA6",
    "rFRC" = "#2AC200",
    "FDC" = "#088B1A"
  )) +
  theme(legend.position = "right",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)+
  xlim(2600, 3100) +
  ylim(4100, 4600) 
dev.off()
```

###Expression
```{r}
fl_df <- fl_df %>%
  mutate(CCR7_order = ifelse(CCR7 > 1, 1, 0)) %>% 
  mutate(CR2_order = ifelse(CR2 > 1, 1, 0)) %>% 
  mutate(THY1_order = ifelse(THY1 > 1, 1, 0)) %>%
  mutate(PDGFRB_order = ifelse(PDGFRB > 1, 1, 0)) %>%
  mutate(CXCL13_order = ifelse(CXCL13 > 1, 1, 0))  %>%
  mutate(CXCR3_order = ifelse(CXCR3 > 1, 1, 0))  %>%
  mutate(CXCL9_order = ifelse(CXCL9 > 1, 1, 0))  %>%
  mutate(CXCL10_order = ifelse(CXCL10 > 1, 1, 0))  %>%
  mutate(CXCL11_order = ifelse(CXCL11 > 1, 1, 0))  %>%
  mutate(CXCL12_order = ifelse(CXCL12 > 1, 1, 0))  %>%
  mutate(CCL21_order = ifelse(CCL21 > 1, 1, 0))
```

```{r}
p2 <- ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CXCL13)) +
  geom_point(data = subset(fl_df, CXCL13_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(fl_df, CXCL13_order == 1),
             size = 0.5, aes(color = CXCL13), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1) +
  xlim(2600, 3100) +
  ylim(4100, 4600)

p3 <- ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CXCL12)) +
  geom_point(data = subset(fl_df, CXCL12_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(fl_df, CXCL12_order == 1),
             size = 0.5, aes(color = CXCL12), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1) +
  xlim(2600, 3100) +
  ylim(4100, 4600)

p4 <- ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CCL21)) +
  geom_point(data = subset(fl_df, CCL21_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(fl_df, CCL21_order == 1),
             size = 0.5, aes(color = CCL21), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1) +
  xlim(2600, 3100) +
  ylim(4100, 4600)

p5 <- ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CXCR3)) +
  geom_point(data = subset(fl_df, CXCR3_order == 0),
             size = 1, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(fl_df, CXCR3_order == 1),
             size = 1, aes(color = CXCR3), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "right",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1) +
  xlim(2600, 3100) +
  ylim(4100, 4600)

p6 <- ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CXCL9)) +
  geom_point(data = subset(fl_df, CXCL9_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(fl_df, CXCL9_order == 1),
             size = 0.5, aes(color = CXCL9), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1) +
  xlim(2600, 3100) +
  ylim(4100, 4600)

```

```{r, fig.height=10, fig.width=10}
pdf(paste0(out_dir_figs, "xenium_fl_expression_zoom_spatial.pdf"), width = 8, height = 8)
cowplot::plot_grid(#p1, p3, p5,
                   p2, p3, p4, p6, 
                   ncol = 2, nrow = 2)
dev.off()
```

##DLBCL
```{r}
dlbcl <- subset(merged_obj, type %in% c("DLBCL"))
genes <- c("CXCL13", "CXCL12", "CCL21", "CCL19", "CXCL9", "CXCL10", "CXCL11", "CCR7", "CR2", "CR1", "PDCD1", "CD8A", "THY1", "CXCR3", "IFNG")
metadata_dlbcl <- as.data.frame(dlbcl@meta.data)
DefaultAssay(dlbcl) <- "RNA"
dlbcl_df <-  cbind(metadata_dlbcl, FetchData(dlbcl, genes))
unique(dlbcl_df$cell_type_binned)
```

###Cell types
```{r}
pdf(paste0(out_dir_figs, "xenium_dlbcl_celltypes_full_spatial.pdf"), width = 4, height = 4)
ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned)) +
  # Plot the background points with #f0f0f0
  geom_point(data = subset(dlbcl_df, cell_type_binned %in% c("BEC", "FL-B", "Myeloid", "T")),
             aes(alpha = 1), size = 0.25, color = "#f0f0f0") +
  
    # Plot the "FL-B" cells with alpha = 0
  geom_point(data = subset(dlbcl_df, cell_type_binned == "DLBCL-B"),
             aes(alpha = 0), size = 0.25, color = "#83CAE8") +
  
  # Plot all other colored points with alpha = 1
  geom_point(data = subset(dlbcl_df, cell_type_binned %in% c("FRC", "rFRC", "FDC")),
             aes(alpha = 1), size = 0.25) +
  
  scale_color_manual(values = c(
    "BEC" = "#f0f0f0",
    "Myeloid" = "#f0f0f0",
    "T-TOX" = "#f0f0f0",
    "T-REG"= "#f0f0f0", 
    "T-FH"= "#f0f0f0", 
    "T-H" = "#f0f0f0",
    "FL-B" = "#f0f0f0",
    "DLBCL-B" = "#83CAE8",
    "FRC" = "#BEDEA6",
    "rFRC" = "#2AC200",
    "FDC" = "#088B1A"))+
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)
dev.off()
```
```{r}
pdf(paste0(out_dir_figs, "xenium_dlbcl_celltypes_full_spatial_legend.pdf"), width = 4, height = 4)
ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned)) +
  # Plot the background points with #f0f0f0
  geom_point(data = subset(dlbcl_df, cell_type_binned %in% c("BEC", "FL-B", "Myeloid", "T")),
             aes(alpha = 1), size = 1, color = "#f0f0f0") +
  
    # Plot the "FL-B" cells with alpha = 0
  geom_point(data = subset(dlbcl_df, cell_type_binned == "DLBCL-B"),
             aes(alpha = 0), size = 1, color = "#83CAE8") +
  
  # Plot all other colored points with alpha = 1
  geom_point(data = subset(dlbcl_df, cell_type_binned %in% c("FRC", "rFRC", "FDC")),
             aes(alpha = 1), size = 1) +
  
  scale_color_manual(values = c(
    "BEC" = "#f0f0f0",
    "Myeloid" = "#f0f0f0",
    "T-TOX" = "#f0f0f0",
    "T-REG"= "#f0f0f0", 
    "T-FH"= "#f0f0f0", 
    "T-H" = "#f0f0f0",
    "FL-B" = "#f0f0f0",
    "DLBCL-B" = "#83CAE8",
    "FRC" = "#BEDEA6",
    "rFRC" = "#2AC200",
    "FDC" = "#088B1A"))+
  theme_bw() +
  theme(legend.position = "right",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)
dev.off()
```

```{r}
pdf(paste0(out_dir_figs, "xenium_dlbcl_celltypes_zoom_spatial.pdf"), width = 4, height = 4)
ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned)) +
  geom_point(data = subset(dlbcl_df, cell_type_binned %in% c("BEC", "FL-B", "Myeloid", "T-TOX", "T-REG", "T-FH", "T-H")),
             aes(alpha = 1), size = 0.5, color = "#f0f0f0") +
  
    # Plot the "FL-B" cells with alpha = 0
  geom_point(data = subset(dlbcl_df, cell_type_binned == "DLBCL-B"),
             aes(alpha = 0.5), size = 0.5) +
  
  # Plot all other colored points with alpha = 1
  geom_point(data = subset(dlbcl_df, cell_type_binned %in% c("FRC", "rFRC", "FDC")),
             aes(alpha = 1), size = 0.5) +
  
  theme_bw() +
  scale_color_manual(values = c(
    "BEC" = "#f0f0f0",
    "Myeloid" = "#f0f0f0",
    "T" = "#f0f0f0",
    "FL-B" = "#f0f0f0",
    "DLBCL-B" = "#83CAE8",
    "FRC" = "#BEDEA6",
    "rFRC" = "#2AC200",
    "FDC" = "#088B1A"
  )) +
  theme(legend.position = "right",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)+
  xlim(800, 1300) +
  ylim(4100, 4600)
dev.off()
```

###Expression
```{r}
dlbcl_df <- dlbcl_df %>%
  mutate(CCR7_order = ifelse(CCR7 > 1, 1, 0)) %>% 
  mutate(CR2_order = ifelse(CR2 > 4, 1, 0)) %>% 
  mutate(THY1_order = ifelse(THY1 > 1, 1, 0)) %>%
  mutate(CXCL13_order = ifelse(CXCL13 > 2, 1, 0))  %>%
  mutate(CXCR3_order = ifelse(CXCR3 > 0, 1, 0))  %>%
  mutate(CXCL9_order = ifelse(CXCL9 > 1, 1, 0))  %>%
  mutate(CD8A_order = ifelse(CD8A > 1, 1, 0))  %>%
  mutate(PDCD1_order = ifelse(PDCD1 > 1, 1, 0))  %>%
  mutate(CXCL10_order = ifelse(CXCL10 > 1, 1, 0))  %>%
  mutate(CXCL11_order = ifelse(CXCL11 > 1, 1, 0))  %>%
  mutate(CXCL12_order = ifelse(CXCL12 > 1, 1, 0))  %>%
  mutate(CCL21_order = ifelse(CCL21 > 1, 1, 0))
```

```{r}
p4 <- ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CCL21)) +
  geom_point(data = subset(dlbcl_df, CCL21_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(dlbcl_df, CCL21_order == 1),
             size = 0.5, aes(color = CCL21), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)+
  xlim(1000, 1500) +
  ylim(4000, 4500)

p3 <- ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CXCL12)) +
  geom_point(data = subset(dlbcl_df, CXCL12_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(dlbcl_df, CXCL12_order == 1),
             size = 0.5, aes(color = CXCL12), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)+
  xlim(1000, 1500) +
  ylim(4000, 4500)
  
p2 <- ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CXCL13)) +
  geom_point(data = subset(dlbcl_df, CXCL13_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(dlbcl_df, CXCL13_order == 1),
             size = 0.5, aes(color = CXCL13), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)+
  xlim(1000, 1500) +
  ylim(4000, 4500)

p6 <- ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = CXCL9)) +
  geom_point(data = subset(dlbcl_df, CXCL9_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 0.7) +
    geom_point(data = subset(dlbcl_df, CXCL9_order == 1),
             size = 0.5, aes(color = CXCL9), alpha = 0.7) +
  scale_color_gradientn(colors = c("#f9f9f9", "#AD2E24")) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)+
  xlim(1000, 1500) +
  ylim(4000, 4500)


```

```{r, fig.height=10, fig.width=10}
pdf(paste0(out_dir_figs, "xenium_dlbcl_expression_zoom_spatial.pdf"), width = 8, height = 8)
cowplot::plot_grid(#p1, p3, p5,
                   p2, p3, p4, p6, 
                   ncol = 2, nrow = 2)
dev.off()
```

##CXCL13 overview (Fig. 2)
```{r}
fl_df <- mutate(fl_df, cell_type_fdc = case_when(
  cell_type_binned %in% c("FDC")~"FDC",
  cell_type_binned %in% c("FL-B", "T-FH", "FRC", "Myeloid", "T-REG", "T-TOX", "T-H", "rFRC", "DLBCL-B", "BEC")~"Other"))
```

```{r}
# Define color mapping for discrete groups
color_map <- c("Other" = "#f0f0f0", "FDC" = "#088B1A")

p1 <- ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y)) +
  
  # Background cells in light grey
  geom_point(data = subset(fl_df, cell_type_fdc == "Other"), 
             aes(color = "Other"), size = 0.5, alpha = 0.1) +
  
  # FDC cells in distinct color
  geom_point(data = subset(fl_df, cell_type_fdc == "FDC"), 
             aes(color = "FDC"), size = 1, alpha = 1) +
  
  # CXCL13 expression using fill
  geom_point(data = subset(fl_df, CXCL13_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 1) +
  
  # CXCL13 expression with color distinction for FDC vs non-FDC cells
  geom_point(data = subset(fl_df, CXCL13_order == 1),
             aes(fill = CXCL13, color = ifelse(cell_type_fdc == "FDC", "FDC", "Other")), 
             size = 1, alpha = 1, shape = 21) +
  
  # Color scale for discrete groups
  scale_color_manual(values = color_map, na.value = NA) + 

  # Fill scale for CXCL13
  scale_fill_gradientn(colors = c("#f9f9f9", "#AD2E24"), guide = "colorbar") +

  theme_bw() +
  theme(legend.position = "right",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)
```

```{r}
dlbcl_df <- mutate(dlbcl_df, cell_type_fdc = case_when(
  cell_type_binned %in% c("FDC")~"FDC",
  #cell_type_binned %in% c("T-TOX")~"T-TOX"
  cell_type_binned %in% c("FL-B", "T-FH", "FRC", "T-TOX", "Myeloid", "T-REG", "T-H", "rFRC", "DLBCL-B", "BEC")~"Other"))
```

```{r}
# Define color mapping for discrete groups
color_map <- c("Other" = "#f9f9f9", 
               #"T-TOX" = "mediumpurple1", 
               "FDC" = "#088B1A")

p2 <- ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y)) +
  
  # Background cells in light grey
  geom_point(data = subset(dlbcl_df, cell_type_fdc == "Other"), 
             aes(color = "Other"), size = 0.5, alpha = 0.1) +
  
  # FDC cells in distinct color
  geom_point(data = subset(dlbcl_df, cell_type_fdc == "FDC"), 
             aes(color = "FDC"), size = 1, alpha = 1) +

  #geom_point(data = subset(dlbcl_df, cell_type_fdc == "T-TOX"), 
  #           aes(color = "T-TOX"), size = 2, alpha = 1) +
  
  # CXCL13 expression using fill
  geom_point(data = subset(dlbcl_df, CXCL13_order == 0),
             size = 0.5, color = "#f9f9f9", alpha = 1) +
  
  # CXCL13 expression with color distinction for FDC vs non-FDC cells
  geom_point(data = subset(dlbcl_df, CXCL13_order == 1),
             aes(fill = CXCL13, color = ifelse(cell_type_fdc %in% "FDC", "FDC", "Other")), 
             size = 1, alpha = 1, shape = 21) +
  
  # Color scale for discrete groups
  scale_color_manual(values = color_map, na.value = NA) + 

  # Fill scale for CXCL13
  scale_fill_gradientn(colors = c("#f9f9f9", "#AD2E24"), guide = "colorbar") +
  theme_bw() +
  theme(legend.position = "right",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)

```
```{r}
pdf(paste0(out_dir_figs, "xenium_fdc_cxcl13_spatial.pdf"), width = 8, height = 4)
cowplot::plot_grid(p1, p2, ncol = 2)
dev.off()
```

##CXCL13+ T cells (not shown)
```{r}
dlbcl_df <- mutate(dlbcl_df, cell_type_ttox = case_when(
  cell_type_binned %in% c("T-TOX")~"T-TOX",
  cell_type_binned %in% c("FL-B", "T-FH", "FRC", "Myeloid", "T-REG", "T-H", "rFRC", "DLBCL-B", "BEC", "FDC")~"Other"))

dlbcl_df <- dlbcl_df %>%
  mutate(CXCL13_order = ifelse(CXCL13 > 3, 1, 0))  %>%
  mutate(CD8A_order = ifelse(CD8A > 1, 1, 0))  %>%
  mutate(PDCD1_order = ifelse(PDCD1 > 1, 1, 0)) 
```

```{r}
# Define color mapping for discrete groups
color_map <- c("Other" = "#f0f0f0", 
               "T-TOX" = "mediumpurple1") 
               #"FDC" = "#088B1A")

ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y)) +

    # Background cells in light grey
  geom_point(data = subset(dlbcl_df, cell_type_ttox == "Other"), 
             aes(color = "Other"), size = 0.5, alpha = 0.1) +
  
    # T cells in distinct color
  geom_point(data = subset(dlbcl_df, cell_type_ttox == "T-TOX"), 
             aes(color = "T-TOX"), size = 1, alpha = 1) +
  
  # CXCL13 expression using fill
 # geom_point(data = subset(dlbcl_df, CXCL13_order == 0),
#             size = 0.5, color = "#f9f9f9", alpha = 1) +
  
  # CXCL13 expression with color distinction for FDC vs non-FDC cells
  #geom_point(data = subset(dlbcl_df, CXCL13_order == 1),
  #           aes(fill = CXCL13, color = ifelse(cell_type_ttox %in% "T-TOX", "T-TOX", "Other")), 
  #           size = 1, alpha = 1, shape = 21) +
  
  # Color scale for discrete groups
  scale_color_manual(values = color_map, na.value = NA) + 

  # Fill scale for CXCL13
  scale_fill_gradientn(colors = c("#f9f9f9", "#AD2E24"), guide = "colorbar") +
  theme_bw() +
  theme(legend.position = "right",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        aspect.ratio = 1)
```

#CXCR3+ T cells
```{r}
# Define CXCR3+ status across all cell types
RidgePlot(subset(merged_obj, cell_type_binned %in% c("T-TOX")), features="CXCR3")
cxcr3_positive <- GetAssayData(merged_obj, assay = "RNA", slot = "data")["CXCR3", ] > 2
merged_obj$CXCR3_status <- ifelse(cxcr3_positive, "CXCR3+", "CXCR3-")

# Define IFNG+ status across all cell types
RidgePlot(subset(merged_obj, cell_type_binned %in% c("T-TOX")), features="IFNG")
ifng_positive <- GetAssayData(merged_obj, assay = "RNA", slot = "data")["IFNG", ] > 2
merged_obj$IFNG_status <- ifelse(ifng_positive, "IFNG+", "IFNG-")

# Define CXCL9+ status across all cell types
RidgePlot(merged_obj, features="CXCL9")
cxcl9_positive <- GetAssayData(merged_obj, assay = "RNA", slot = "data")["CXCL9", ] > 2
merged_obj$CXCL9_status <- ifelse(cxcl9_positive, "CXCL9+", "CXCL9-")
```

##Tidy data
```{r}
merged_obj@meta.data <- mutate(merged_obj@meta.data, cell_type_binned_cxcr3 = case_when(cell_type_binned %in% c("T-TOX", "T-H", "T-FH", "T-REG") & CXCR3_status == "CXCR3+"~"CXCR3+ T",
                                                      cell_type_binned %in% c("T-TOX", "T-H", "T-FH", "T-REG") & CXCR3_status == "CXCR3-"~"CXCR3- T",
                                                      cell_type_binned %in% c("rFRC", "FRC") & CXCL9_status == "CXCL9+"~"CXCL9+ FRC",
                                                      cell_type_binned %in% c("rFRC", "FRC") & CXCL9_status == "CXCL9-"~"CXCL9- FRC",
                                                      cell_type_binned %in% c("FL-B", "Myeloid", "FDC", "BEC", "DLBCL-B")~"Other"))
unique(merged_obj@meta.data$cell_type_binned_cxcr3)

merged_obj@meta.data <- mutate(merged_obj@meta.data, cell_type_binned_ifng = case_when(cell_type_binned %in% c("T-TOX", "T-H", "T-FH", "T-REG") & IFNG_status == "IFNG+"~"IFNG+ T",
                                                      cell_type_binned %in% c("T-TOX", "T-H", "T-FH", "T-REG") & IFNG_status == "IFNG-"~"IFNG- T",
                                                      cell_type_binned %in% c("rFRC", "FRC") & CXCL9_status == "CXCL9+"~"CXCL9+ FRC",
                                                      cell_type_binned %in% c("rFRC", "FRC") & CXCL9_status == "CXCL9-"~"CXCL9- FRC",
                                                      cell_type_binned %in% c("FL-B", "Myeloid", "FDC", "BEC", "DLBCL-B")~"Other"))
unique(merged_obj@meta.data$cell_type_binned_ifng)
```

```{r}
genes <- c("CCL21", "CCL19", "CXCL9", "CXCL10", "CXCL11", "CCR7", "CXCR3", "IFNG")
metadata <- as.data.frame(merged_obj@meta.data)
DefaultAssay(merged_obj) <- "RNA"
merged_df <-  cbind(metadata, FetchData(merged_obj, genes))

fl_df <- subset(merged_df, type %in% c("FL"))
dlbcl_df <- subset(merged_df, type %in% c("DLBCL"))
```

##CXCL9-CXCR3 spatial plots
```{r}
p1 <- ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned_cxcr3)) +
  geom_point(data = subset(fl_df, cell_type_binned_cxcr3 %in% c("Other", "CXCR3- T", "CXCL9- FRC")),
             aes(alpha = 1), size = 0.25, color = "#f0f0f0") +
  
  geom_point(data = subset(fl_df, cell_type_binned_cxcr3 == "CXCR3+ T"),
             aes(alpha = 1), size = 1, color = "mediumpurple1") +

  geom_point(data = subset(fl_df, cell_type_binned_cxcr3 == "CXCL9+ FRC"),
             aes(alpha = 1), size = 1, color = "#2AC200") +

  #scale_color_manual(values = c(
  #  "Other" = "#f0f0f0",
  #  "CXCR3- T" = "#f0f0f0",
  #  "CXCR3+ T" = "mediumpurple1"))+
  theme_bw() +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank()
)
```

```{r}
p2 <- ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned_cxcr3)) +
  geom_point(data = subset(dlbcl_df, cell_type_binned_cxcr3 %in% c("Other", "CXCR3- T", "CXCL9- FRC")),
             aes(alpha = 1), size = 0.25, color = "#f0f0f0") +
  
  geom_point(data = subset(dlbcl_df, cell_type_binned_cxcr3 == "CXCR3+ T"),
             aes(alpha = 1), size = 1, color = "mediumpurple1") +

  geom_point(data = subset(dlbcl_df, cell_type_binned_cxcr3 == "CXCL9+ FRC"),
             aes(alpha = 1), size = 1, color = "#2AC200") +
  #scale_color_manual(values = c(
  #  "Other" = "#f0f0f0",
  #  "CXCR3- T" = "#f0f0f0",
  #  "CXCR3+ T" = "mediumpurple1"))+
  theme_bw() +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank()
)
```

##Density
```{r}
library(spatstat)
```

###FL
```{r}
cxcl9_pos_rfrc <- fl_df %>% filter(cell_type_binned_cxcr3 == "CXCL9+ FRC")
cxcl9_neg_rfrc <- fl_df %>% filter(cell_type_binned_cxcr3 == "CXCL9- FRC")
cxcr3_pos_t <- fl_df %>% filter(cell_type_binned_cxcr3 == "CXCR3+ T")
#cxcr3_neg_t <- fl_df %>% filter(cell_type_binned_cxcr3 == "CXCR3- T")
```

```{r}
# Convert coordinates into spatial point patterns (ppp format)
buffer <- 100  # Increase buffer slightly

x_min <- min(fl_df$nucleus_centroid_x) - buffer
x_max <- max(fl_df$nucleus_centroid_x) + buffer
y_min <- min(fl_df$nucleus_centroid_y) - buffer
y_max <- max(fl_df$nucleus_centroid_y) + buffer

window <- owin(xrange = c(x_min, x_max), yrange = c(y_min, y_max))
print(window)

# Convert cell positions to spatial point patterns
cxcr3_pos_t_ppp <- ppp(cxcr3_pos_t$nucleus_centroid_x, cxcr3_pos_t$nucleus_centroid_y, window = window)
cxcl9_pos_rfrc_ppp <- ppp(cxcl9_pos_rfrc$nucleus_centroid_x, cxcl9_pos_rfrc$nucleus_centroid_y, window = window)
cxcl9_neg_rfrc_ppp <- ppp(cxcl9_neg_rfrc$nucleus_centroid_x, cxcl9_neg_rfrc$nucleus_centroid_y, window = window)

# Perform Kernel Density Estimation (KDE) for CXCR3+ T cells
sigma_value <- 50  # Adjust as needed
cxcr3_pos_density <- density(cxcr3_pos_t_ppp, sigma = sigma_value)

# Extract CXCR3+ T cell density at CXCL9+ and CXCL9- rFRC locations
cxcr3_pos_at_cxcl9_pos_rfrc <- as.data.frame(spatstat.geom::interp.im(cxcr3_pos_density, cxcl9_pos_rfrc_ppp))
cxcr3_pos_at_cxcl9_neg_rfrc <- as.data.frame(spatstat.geom::interp.im(cxcr3_pos_density, cxcl9_neg_rfrc_ppp))

colnames(cxcr3_pos_at_cxcl9_pos_rfrc) <- "density"
colnames(cxcr3_pos_at_cxcl9_neg_rfrc) <- "density"

# Create a combined dataframe for both CXCL9+ FRC and CXCL9- FRC for comparison
density_df <- rbind(
  data.frame(density = cxcr3_pos_at_cxcl9_pos_rfrc$density, group = "CXCL9+ FRC"),
  data.frame(density = cxcr3_pos_at_cxcl9_neg_rfrc$density, group = "CXCL9- FRC")
)

# Plot the comparison between CXCL9+ and CXCL9- FRCs
density_df$group <- factor(density_df$group, levels = c("CXCL9+ FRC", "CXCL9- FRC"))
p3 <- ggplot(density_df, aes(x = group, y = density)) +
  geom_boxplot(aes(fill = group), outlier.shape = NA, width = 0.5, alpha = 0.7) +
  #geom_jitter(shape = 16, position = position_jitter(0.2), alpha = 0.4) +
  scale_fill_manual(values = c("#2AC200", "#f0f0f0")) + # Adjust the colors
  ylab("CXCR3+ T Density")+
  coord_flip()+
  stat_compare_means(method = "wilcox.test", label = "p.signif", comp = list(c("CXCL9+ FRC", "CXCL9- FRC"))) +
  theme_bw() +
  theme(
    plot.title = element_blank(),
    plot.margin = margin(10, 10, 10, 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(color = "black"),
    legend.position = "none",
    aspect.ratio = 0.5,
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.x = element_text(size = 12, color = "black"),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = 12, color = "black")
  )
```

###DLBCL
```{r}
cxcl9_pos_rfrc <- dlbcl_df %>% filter(cell_type_binned_cxcr3 == "CXCL9+ FRC")
cxcl9_neg_rfrc <- dlbcl_df %>% filter(cell_type_binned_cxcr3 == "CXCL9- FRC")
cxcr3_pos_t <- dlbcl_df %>% filter(cell_type_binned_cxcr3 == "CXCR3+ T")
#cxcr3_neg_t <- dlbcl_df %>% filter(cell_type_binned_cxcr3 == "CXCR3- T")
```

```{r}
# Convert coordinates into spatial point patterns (ppp format)
buffer <- 100  # Increase buffer slightly

x_min <- min(dlbcl_df$nucleus_centroid_x) - buffer
x_max <- max(dlbcl_df$nucleus_centroid_x) + buffer
y_min <- min(dlbcl_df$nucleus_centroid_y) - buffer
y_max <- max(dlbcl_df$nucleus_centroid_y) + buffer

window <- owin(xrange = c(x_min, x_max), yrange = c(y_min, y_max))
print(window)

# Convert cell positions to spatial point patterns
cxcr3_pos_t_ppp <- ppp(cxcr3_pos_t$nucleus_centroid_x, cxcr3_pos_t$nucleus_centroid_y, window = window)
cxcl9_pos_rfrc_ppp <- ppp(cxcl9_pos_rfrc$nucleus_centroid_x, cxcl9_pos_rfrc$nucleus_centroid_y, window = window)
cxcl9_neg_rfrc_ppp <- ppp(cxcl9_neg_rfrc$nucleus_centroid_x, cxcl9_neg_rfrc$nucleus_centroid_y, window = window)

# Perform Kernel Density Estimation (KDE) for CXCR3+ T cells
sigma_value <- 50  # Adjust as needed
cxcr3_pos_density <- density(cxcr3_pos_t_ppp, sigma = sigma_value)

# Extract CXCR3+ T cell density at CXCL9+ and CXCL9- rFRC locations
cxcr3_pos_at_cxcl9_pos_rfrc <- as.data.frame(spatstat.geom::interp.im(cxcr3_pos_density, cxcl9_pos_rfrc_ppp))
cxcr3_pos_at_cxcl9_neg_rfrc <- as.data.frame(spatstat.geom::interp.im(cxcr3_pos_density, cxcl9_neg_rfrc_ppp))

colnames(cxcr3_pos_at_cxcl9_pos_rfrc) <- "density"
colnames(cxcr3_pos_at_cxcl9_neg_rfrc) <- "density"

# Create a combined dataframe for both CXCL9+ FRC and CXCL9- FRC for comparison
density_df <- rbind(
  data.frame(density = cxcr3_pos_at_cxcl9_pos_rfrc$density, group = "CXCL9+ FRC"),
  data.frame(density = cxcr3_pos_at_cxcl9_neg_rfrc$density, group = "CXCL9- FRC")
)

# Plot the comparison between CXCL9+ and CXCL9- FRCs
density_df$group <- factor(density_df$group, levels = c("CXCL9+ FRC", "CXCL9- FRC"))
p4 <- ggplot(density_df, aes(x = group, y = density)) +
  geom_boxplot(aes(fill = group), outlier.shape = NA, width = 0.5, alpha = 0.7) +
  #geom_jitter(shape = 16, position = position_jitter(0.2), alpha = 0.4) +
  scale_fill_manual(values = c("#2AC200", "#f0f0f0")) + # Adjust the colors
  ylab("CXCR3+ T Density")+
  coord_flip()+
  stat_compare_means(method = "wilcox.test", label = "p.signif", comp = list(c("CXCL9+ FRC", "CXCL9- FRC"))) +
  theme_bw() +
  theme(
    plot.title = element_blank(),
    plot.margin = margin(10, 10, 10, 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(color = "black"),
    legend.position = "none",
    aspect.ratio = 0.5,
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 12, color = "black"),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = 12, color = "black")
  )
```

###Save plots
```{r}
pdf(paste0(out_dir_figs, "xenium_cxcr3t_cxcl9frc_spatial.pdf"), width = 6, height = 3)
cowplot::plot_grid(p1, p2, ncol = 2)
dev.off()

pdf(paste0(out_dir_figs, "xenium_cxcr3t_cxcl9frc_box.pdf"), width = 8, height = 2)
cowplot::plot_grid(p3, p4, ncol = 2)
dev.off()
```
