---
title: "Figure3"
output: html_document
date: "2024-02-29"
---

# Load pacakges
```{r setup, include=FALSE}
library(Seurat)
library(data.table)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(tidyverse)
library(UpSetR)
library(circlize)
library(ComplexHeatmap)
```

# Load objects
```{r setup, include=FALSE}
scRNA <- readRDS("") #Available in zenodo as scRNA_LN_stromal_cells.rds
source("/path/to/Figure_config.R")
```

## B
## SCENIC output can be directly read targene_df from Zenodo as GRN_90pt.rds
```{r}
#  SCENIC output-target genes are assigned to each TF
# Note: This illustrates how to process SCENIC output for network re-construction (i.e. creating GRN_90pt.rds from SCENIC output)
reg_files <- list.files("path/to/regulon", 
                       pattern = ".*\\(\\+\\)\\.tsv$", 
                       full.names = T)

df_list <- list()
for (file in reg_files) {
  # the regex matches any characters except "/" that are right before a "(+).tsv" and thereby fetches the TF-names
  TF_name <- str_extract(file, "[^\\/]+(?=\\(\\+\\)\\.tsv)")
  regulon_df <- read.csv(file, sep = "\t", header = F, col.names = c("target", "count"))
  regulon_df <- mutate(regulon_df, TF = TF_name) 
  df_list[[TF_name]] <- regulon_df
}
# targene_df_raw contains all target genes for the TFs, even the ones with
# counts < 80 that were not used for the activity calculation
empty_df <- data.frame(TF = character(), target = character(),  count = numeric())
targene_df_raw <- purrr::reduce(df_list, bind_rows, .init = empty_df)
# make another copy with only the target genes that were used for the activity calculation (observed in >80% of runs; set in min-regulon-gene-occurrence in config.yaml
targene_df <- filter(targene_df_raw, count > 90) # This is the GRN used in C
```

## B
```{r}
targene_df <- readRDS("") #Available in Zenodo as GRN_90pt.rds, you can read directly this and skip the part above

#Read in differentially expressed genes
#Output from chunk of code from Figure 2 panel B but data also available as source data in the manuscript
#DE_all <- FindMarkers(tmp_obj, ident.2 = "rLN", ident.1 = "DLBCL", pseudocount.use = 0.01, test.use = "MAST", only.pos = F, logfc.threshold = 0, assay = "RNA")

de_genes_scenic <- list()
for (g in c("BEC", "LEC", "FRC")){
  de_genes <- read.table(paste0("/g/scb2/zaugg/amathiou/2021July_LNS012/3.DiffExpression/20220901_Abe_Gate/", g, "_DEgenes_MAST_dlbclVSrln.tsv"))
  de_genes_scenic[[paste0(g, " DLBCL")]] <- rownames(de_genes[which(de_genes$p_val_adj < 0.05 & de_genes$avg_log2FC > 1), ])
  de_genes_scenic[[paste0(g, " rLN")]] <- rownames(de_genes[which(de_genes$p_val_adj < 0.05 & de_genes$avg_log2FC < -1), ])
}
pop_de <- list()
pop_de[["BEC"]] <- unique(unlist(de_genes_scenic[c("BEC DLBCL", "BEC rLN")]))
pop_de[["LEC"]] <- unique(unlist(de_genes_scenic[c("LEC DLBCL", "LEC rLN")]))
pop_de[["FRC"]] <- unique(unlist(de_genes_scenic[c("FRC DLBCL", "FRC rLN")]))

#This upset plot is part of supplementary figure8 
upset(fromList(de_genes_scenic), nsets = 15, order.by = "freq", point.size = 1.5)

#
sig_TFs <- list()
fisher_df1 <- list()
fisher_df2 <- list()
for (p in c("BEC", "LEC", "FRC")){
  tmp_genes <- intersect(pop_de[[p]], unique(targene_df$target)) #Take those that are in the network
  tmp_targene_df <- targene_df[which(targene_df$target %in% tmp_genes), ] #Lets look only at all the DE genes, both conditions
  dlbcl_genes <- intersect(tmp_genes, de_genes_scenic[[paste0(p, " DLBCL")]])
  rln_genes <- intersect(tmp_genes, de_genes_scenic[[paste0(p, " rLN")]])
  count_tfs_genes <- data.frame(TF = unique(tmp_targene_df[which(tmp_targene_df$target %in% tmp_genes), "TF"])) #how many of the condition genes these TFs target
  rownames(count_tfs_genes) <- count_tfs_genes$TF
  
  table_tfs <- data.frame() #Fisher test input
  for (t in rownames(count_tfs_genes)){
    #A bit repeatitive with the cont table, but i like the visualization of this table
    tf_targets <- tmp_targene_df[which(tmp_targene_df$TF == t & tmp_targene_df$target %in% tmp_genes), "target"]
    count_tfs_genes[t, "DLBCL_TF"] <- length(intersect(tf_targets, dlbcl_genes)) + 1
    count_tfs_genes[t, "rLN_TF"] <- length(intersect(tf_targets, rln_genes)) + 1
    count_tfs_genes[t, "DLBCL_rest"] <- length(setdiff(dlbcl_genes, tf_targets)) + 1
    count_tfs_genes[t, "rLN_rest"] <- length(setdiff(rln_genes, tf_targets)) + 1
    #Contigency table
    cont_table <- data.frame()
    cont_table['DLBCL', t] <- count_tfs_genes[t, "DLBCL_TF"]
    cont_table['DLBCL', 'rest'] <- count_tfs_genes[t, "DLBCL_rest"]
    cont_table['rLN', t] <- count_tfs_genes[t, "rLN_TF"]
    cont_table['rLN', 'rest'] <- count_tfs_genes[t, "rLN_rest"]
    table_tfs[t, 'TF'] <- t
    table_tfs[t, 'Fisher_pvalue'] <- fisher.test(cont_table, alternative='two.sided', conf.int = TRUE)$p.value
    table_tfs[t, 'OR'] <- fisher.test(cont_table, alternative='two.sided', conf.int = TRUE)$estimate
  }
  table_tfs$fdr <- p.adjust(table_tfs$Fisher_pvalue, method="fdr")
  table_tfs$padj <- ifelse(table_tfs$fdr < 0.05, 
                          "<0.05", 
                          "n.s.")

  table_tfs$log2OR <- log2(table_tfs$OR)
  table_tfs <- table_tfs[order(table_tfs$log2OR), ]
  levels_plot <- rownames(table_tfs)

  #Plot enrichment
  q3 <- ggplot(table_tfs, aes(x = factor(TF, levels = levels_plot), y = log2OR, fill = padj)) + 
      geom_bar(stat = "identity") + 
      theme_bw() + 
      theme(text = element_text(size = 15), 
                     axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
            legend.position = "right") + xlab("TFs") + 
    ggtitle(p) + 
    scale_fill_manual("padj", values = c("n.s." = "#ffe7b1", "<0.05" = "#75cdc8")) # + scale_fill_gradientn(colors = head(viridis(10), 9)
  # These barplots are part of supplemental figure 8
  plot(q3)

  #This will be used to actually constract the panel B
  sig_TFs[[paste0(p, " DLBCL")]] <- table_tfs[which(table_tfs$padj < 0.05 & table_tfs$log2OR > 0), "TF"]
  sig_TFs[[paste0(p, " rLN")]] <- table_tfs[which(table_tfs$padj < 0.05 & table_tfs$log2OR < 0), "TF"]
}

#This upset plot is part of supplementary figure 8
upset(fromList(sig_TFs), nsets = 15, order.by = "freq", point.size = 1.5)


# Start of panel B code
msc_or <- read.table(paste0("/path/to/FisherTF_targets_FRC.tsv"))

plot_tfs <- unique(unlist(sig_TFs))
#tile plot but with log2fc
or_df <- data.frame(Gene = plot_tfs, 
                    BEC = NA,
                    LEC = NA, 
                    MSC = NA)
rownames(or_df) <- or_df$Gene
or_df$Gene <- NULL

for (g in rownames(or_df)){
  if (g %in% c(sig_TFs$`BEC DLBCL`, sig_TFs$`BEC rLN`)){
    or_df[g, "BEC"] <- log2(bec_or[g, "OR"])
  }
  if(g %in% c(sig_TFs$`LEC DLBCL`, sig_TFs$`LEC rLN`)){
    or_df[g, "LEC"] <- log2(lec_or[g, "OR"])
  }
  if(g %in% c(sig_TFs$`Mesenchymal DLBCL`, sig_TFs$`Mesenchymal rLN`)){
    or_df[g, "MSC"] <- log2(msc_or[g, "OR"])
  }
}
col_fun = colorRamp2(c(min(or_df), 0, max(or_df)), c("#8D60B5", "white", "#67B86E"))

Heatmap(or_df, col = col_fun, name = "log2OR") #panel B, not that from this we show in the MS only the FRCs (i.e. Mesenchymal in the object)
```

# Panels E, F, G
#For TCR seq related analysis, see https://github.com/ZauggGroup/LNarch/main/Preprocessing/scTCR_reanalysis_revisions.Rmd
## Object also used for TCR clonal expansion analysis in Figure 3
## Dataset from Roider et al. Nat Cell Biol 2024
## 5' TCR seq single cell object projected to 3' scRNA-seq dataset similarly to the original publication
## Analysis was focused on rLN, FL and DLBCL

```{r Panel E left}
DF_TCRrep <- readRDS("/202402_TCRrepertoire.rds") #Available in zenodo: Tcell clonal information 

DF_5prime_umap <- DF_TCRrep %>% 
  dplyr::select(refUMAP_1, refUMAP_2, predicted.celltype, Entity, PatientID.x, ifng.x) %>% 
  distinct() %>% 
  mutate(Entity=factor(Entity, levels=c("rLN", "FL", "DLBCL")))

ggplot() + 
  geom_point(data=DF_5prime_umap %>% 
               arrange(ifng), aes(x=refUMAP_1, y=refUMAP_2, fill = ifng), size=1.5, stroke=0, shape=21) + 
  scale_color_gradientn(colors = parula(5)) + 
  scale_fill_gradientn(colors = parula(5)) + 
  mytheme_1 +
  theme(legend.position = "top",
        legend.text = element_text(size=8),
        legend.background = element_rect(fill=NA),
        legend.box = unit("cm", x = 0.01),
        legend.spacing.x = unit("cm", x = 0.05),
        legend.spacing.y = unit("cm", x = 0.001),
        panel.border = element_rect(size=0.25, color="black"),
        legend.box.spacing = unit(0, units = "cm"),
        strip.background = element_rect(size=0.25),
        legend.box.margin = unit(c(0,-12.5,-0.1,0), units = "cm"))
```

```{r Panel E right}
set.seed(1)
df_clonotypes <- 
  DF_TCRrep %>% 
  filter(!is.na(raw_clonotype_id)) %>% 
  filter(!is.na(Entity)) %>% 
  dplyr::select(Barcode_full, PatientID.x, raw_clonotype_id, refUMAP_1, refUMAP_2, predicted.celltype, Entity) %>% #missing entity
  distinct() %>% 
  add_count(PatientID.x, raw_clonotype_id, predicted.celltype) %>% ungroup() %>% 
  dplyr::select(-Barcode_full) %>% 
  group_by(PatientID.x, raw_clonotype_id, predicted.celltype) %>%
  summarise(refUMAP_1=median(refUMAP_1), refUMAP_2=median(refUMAP_2), n, Entity, .groups = "drop") %>% 
  distinct() %>% 
  mutate(n=ifelse(n>50, 50, n)) %>% 
  mutate(Entity=factor(Entity, levels=c("rLN", "DLBCL", "FL"))) %>% 
  filter(predicted.celltype %in% c(3, 5, 16))

p1 <- ggplot()+
  geom_point_rast(data=DF_5prime_umap, aes(x=refUMAP_1, y=refUMAP_2), 
                  size=1.5, stroke=0, shape=21,  fill="lightgrey")+
  geom_point_rast(data=DF_5prime_umap, aes(x=refUMAP_1, y=refUMAP_2), size=0.2, 
                  alpha=1, 
                  stroke=0, shape=21)+
  geom_point(data=df_clonotypes %>% 
               filter(n > 2), 
             aes(x=refUMAP_1, y=refUMAP_2, size=n, color=Entity), shape=21, stroke=0.7, alpha=1)+
  scale_color_manual(values = c("FL" = "#8BCDE0", "rLN" = "#8D60B5", "DLBCL" = "#67B86E"), guide="none")+
  scale_fill_manual(values = c("FL" = "#8BCDE0", "rLN" = "#8D60B5", "DLBCL" = "#67B86E"), guide="none")+
  scale_size_continuous(range=c(1, 10), 
                        limits=c(3, 50), 
                        breaks=c(3, 20, 35, 50),
                        labels=c("3", "20", "35", "> 50"), 
                        name = NULL)+
  #facet_wrap(~Entity, nrow = 1)+
  geom_text(data = df_clonotypes %>% dplyr::select(PatientID.x, Entity) %>% 
              distinct() %>% 
              add_count(Entity), aes(label = paste0("n = ", n)),
            x = 9.65, y = -6.9, hjust=1, check_overlap = T, size=2.25)+
  xlab("refUMAP-1")+
  ylab("refUMAP-2")+
  mytheme_1+
  theme(legend.position = "top",
        legend.text = element_text(size=8),
        legend.background = element_rect(fill=NA),
        legend.box = unit("cm", x = 0.01),
        legend.spacing.x = unit("cm", x = 0.05),
        legend.spacing.y = unit("cm", x = 0.001),
        panel.border = element_rect(size=0.25, color="black"),
        legend.box.spacing = unit(0, units = "cm"),
        strip.background = element_rect(size=0.25),
        legend.box.margin = unit(c(0,-12.5,-0.1,0), units = "cm"))
plot(p1)
```

```{r Panel F}
DF_TCRrep_cd8 <- DF_TCRrep[which(DF_TCRrep$predicted.celltype %in% c(3, 5, 16)), ]
clonotype_df <- as.data.frame.matrix(table(DF_TCRrep_cd8$raw_clonotype_id, DF_TCRrep_cd8$Entity))
clonotype_freq <- as.data.frame(rowSums(clonotype_df))
colnames(clonotype_freq) <- "Freq"
quantile(clonotype_freq$Freq)

#Panel below is for Supplemental Figure 9 - It basically helps us decide on the threshold
ggplot(clonotype_freq, aes(x = log10(Freq) )) + 
  geom_density() + 
  theme_bw() + 
  #geom_vline(xintercept = log10(quantile(clonotype_freq$Freq, .75, names = F)), 
  #           linetype="dashed", color = "#F54748", size=1) +
  geom_vline(xintercept = log10(4), 
             linetype="dashed", color = "#7F0000", size=1) +
  xlab("log10(Freq_Clone)") + 
  ggtitle("2,558 clonotypes")
#Based on this, we decide that when a clone size equal or greater to 4 it is considered clonally expanded2
```

```{r Prep for panel G, fig.width=5, fig.height=4}
df_clonotypes <- 
  DF_TCRrep_cd8 %>% 
  filter(!is.na(raw_clonotype_id)) %>% 
  filter(!is.na(Entity)) %>% 
  mutate(Entity=factor(Entity, levels=c("rLN", "FL", "DLBCL")))

clone_df <- data.frame(Clone = unique(df_clonotypes$raw_clonotype_id))
clone_df$IFNG <- unlist(lapply(unique(df_clonotypes$raw_clonotype_id), function(clonotype){
  mean(df_clonotypes[which(df_clonotypes$raw_clonotype_id == clonotype), "IFNG"])
  }))
clone_df$Freq <- unlist(lapply(unique(df_clonotypes$raw_clonotype_id), 
                               function(clonotype){
  nrow(df_clonotypes[which(df_clonotypes$raw_clonotype_id == clonotype), ])
  }))

clone_df$Expanded <- ifelse(clone_df$Freq >= 4, 
                            "Expanded", 
                            "Non-expanded")

# SuppFig9
ggplot(clone_df, aes(x = log10(Freq), y = IFNG, color = Expanded, fill = Expanded)) + 
  geom_point() + 
  theme_bw() + 
  ggtitle("CD8+ T cells") + 
  xlab("log10(Clone Size)") + 
  geom_smooth(method = "lm", se = TRUE, fullrange = F) + 
  scale_color_manual(values = c("#7F0000", "darkgrey")) + 
  scale_fill_manual(values = c("#7F0000", "darkgrey"))

clones_meta <- clone_df[, c("Clone", "Expanded")]
colnames(clones_meta) <- c("raw_clonotype_id", "Expanded")
df_clonotypes <- merge(df_clonotypes, clones_meta, by = "raw_clonotype_id")
```

```{r, Panel G left}
clone_entity_bar <- as.data.frame(table(df_clonotypes$Expanded, df_clonotypes$Entity))
colnames(clone_entity_bar) <- c("Expanded", "Entity", "Freq")

ggplot(clone_entity_bar %>% mutate(Entity=factor(Entity, levels=c("rLN", "FL", "DLBCL"))), 
       aes(y = Freq, x = Entity, label=Freq, color = Expanded, fill = Expanded)) + 
  geom_bar(stat = "identity", position = "fill", alpha = 0.5) + 
  scale_color_manual(values = c("#7F0000", "darkgrey")) + 
  scale_fill_manual(values = c("#B13A2C", "lightgrey")) + theme_classic()
```

```{r, Panel G right}
ggplot(df_clonotypes, aes(x = Expanded, y = IFNG)) + 
  geom_boxplot(aes(color = Expanded, fill = Expanded), alpha = 0.5, outlier.size = 0.05) + 
  theme_bw() + 
  facet_wrap(~Entity) + 
  scale_color_manual(values = c("darkgrey", "#7F0000")) +   
  scale_fill_manual(values = c("lightgrey", "#B13A2C")) + stat_compare_means()
```

# Panel J
```{r}
## Load data
xenium_seurat <- readRDS("") #Available in zenodo as "xenium_merged_seurat.rds"

## Define CXCR3+ status across all cell types
RidgePlot(subset(xenium_seurat, cell_type_binned %in% c("T-TOX")), features="CXCR3")
cxcr3_positive <- GetAssayData(xenium_seurat, assay = "RNA", slot = "data")["CXCR3", ] > 2
merged_obj$CXCR3_status <- ifelse(cxcr3_positive, "CXCR3+", "CXCR3-")

## Define CXCL9+ status across all cell types
RidgePlot(xenium_seurat, features="CXCL9")
cxcl9_positive <- GetAssayData(xenium_seurat, assay = "RNA", slot = "data")["CXCL9", ] > 2
merged_obj$CXCL9_status <- ifelse(cxcl9_positive, "CXCL9+", "CXCL9-")

## Tidy data
xenium_seurat@meta.data <- mutate(xenium_seurat@meta.data, cell_type_binned_cxcr3 = case_when(cell_type_binned %in% c("T-TOX", "T-H", "T-FH", "T-REG") & CXCR3_status == "CXCR3+"~"CXCR3+ T",
                                                      cell_type_binned %in% c("T-TOX", "T-H", "T-FH", "T-REG") & CXCR3_status == "CXCR3-"~"CXCR3- T",
                                                      cell_type_binned %in% c("rFRC", "FRC") & CXCL9_status == "CXCL9+"~"CXCL9+ FRC",
                                                      cell_type_binned %in% c("rFRC", "FRC") & CXCL9_status == "CXCL9-"~"CXCL9- FRC",
                                                      cell_type_binned %in% c("FL-B", "Myeloid", "FDC", "BEC", "DLBCL-B")~"Other"))

genes <- c("CXCL9", "CXCR3")
metadata <- as.data.frame(xenium_seurat@meta.data)
DefaultAssay(xenium_seurat) <- "RNA"
merged_df <-  cbind(metadata, FetchData(xenium_seurat, genes))

fl_df <- subset(xenium_seurat, type %in% c("FL"))
dlbcl_df <- subset(xenium_seurat, type %in% c("DLBCL"))

# CXCL9-CXCR3 spatial plots

panel_j_upperleft <- ggplot(fl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned_cxcr3)) +
  geom_point(data = subset(fl_df, cell_type_binned_cxcr3 %in% c("Other", "CXCR3- T", "CXCL9- FRC")),
             alpha = 1, size = 0.25, color = "#ececec", shape = 16) +
  
  geom_point(data = subset(fl_df, cell_type_binned_cxcr3 == "CXCR3+ T"),
             alpha = 0.5, size = 1, color = "mediumpurple1", shape = 16) +

  geom_point(data = subset(fl_df, cell_type_binned_cxcr3 == "CXCL9+ FRC"),
             alpha = 0.5, size = 1, color = "#2AC200", shape = 16) +
  #scale_color_manual(values = c(
  #  "Other" = "#f0f0f0",
  #  "CXCR3- T" = "#f0f0f0",
  #  "CXCR3+ T" = "mediumpurple1"))+
  theme_bw() +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank()
)

panel_j_upperright <- ggplot(dlbcl_df, aes(x = nucleus_centroid_x, y = nucleus_centroid_y, color = cell_type_binned_cxcr3)) +
  geom_point(data = subset(dlbcl_df, cell_type_binned_cxcr3 %in% c("Other", "CXCR3- T", "CXCL9- FRC")),
             alpha = 1, size = 0.25, color = "#ececec", shape = 16) +
  
  geom_point(data = subset(dlbcl_df, cell_type_binned_cxcr3 == "CXCR3+ T"),
             alpha = 0.5, size = 1, color = "mediumpurple1", shape = 16) +

  geom_point(data = subset(dlbcl_df, cell_type_binned_cxcr3 == "CXCL9+ FRC"),
             alpha = 0.5, size = 1, color = "#2AC200", shape = 16) +
  #scale_color_manual(values = c(
  #  "Other" = "#f0f0f0",
  #  "CXCR3- T" = "#f0f0f0",
  #  "CXCR3+ T" = "mediumpurple1"))+
  theme_bw() +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_rect(color = "black"),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank()
)

panel_j_upperleft + panel_j_upperright

## Density of CXCR3+ T cells around CXCL9+ FRCs
### FL
cxcl9_pos_rfrc <- fl_df %>% filter(cell_type_binned_cxcr3 == "CXCL9+ FRC")
cxcl9_neg_rfrc <- fl_df %>% filter(cell_type_binned_cxcr3 == "CXCL9- FRC")
cxcr3_pos_t <- fl_df %>% filter(cell_type_binned_cxcr3 == "CXCR3+ T")

#### Convert coordinates into spatial point patterns (ppp format)
buffer <- 100  # Increase buffer slightly

x_min <- min(fl_df$nucleus_centroid_x) - buffer
x_max <- max(fl_df$nucleus_centroid_x) + buffer
y_min <- min(fl_df$nucleus_centroid_y) - buffer
y_max <- max(fl_df$nucleus_centroid_y) + buffer

window <- owin(xrange = c(x_min, x_max), yrange = c(y_min, y_max))
print(window)

#### Convert cell positions to spatial point patterns
cxcr3_pos_t_ppp <- ppp(cxcr3_pos_t$nucleus_centroid_x, cxcr3_pos_t$nucleus_centroid_y, window = window)
cxcl9_pos_rfrc_ppp <- ppp(cxcl9_pos_rfrc$nucleus_centroid_x, cxcl9_pos_rfrc$nucleus_centroid_y, window = window)
cxcl9_neg_rfrc_ppp <- ppp(cxcl9_neg_rfrc$nucleus_centroid_x, cxcl9_neg_rfrc$nucleus_centroid_y, window = window)

#### Perform Kernel Density Estimation (KDE) for CXCR3+ T cells
sigma_value <- 50  # Adjust as needed
cxcr3_pos_density <- density(cxcr3_pos_t_ppp, sigma = sigma_value)

#### Extract CXCR3+ T cell density at CXCL9+ and CXCL9- rFRC locations
cxcr3_pos_at_cxcl9_pos_rfrc <- as.data.frame(spatstat.geom::interp.im(cxcr3_pos_density, cxcl9_pos_rfrc_ppp))
cxcr3_pos_at_cxcl9_neg_rfrc <- as.data.frame(spatstat.geom::interp.im(cxcr3_pos_density, cxcl9_neg_rfrc_ppp))

colnames(cxcr3_pos_at_cxcl9_pos_rfrc) <- "density"
colnames(cxcr3_pos_at_cxcl9_neg_rfrc) <- "density"

#### Create a combined dataframe for both CXCL9+ FRC and CXCL9- FRC for comparison
density_df_fl <- rbind(
  data.frame(density = cxcr3_pos_at_cxcl9_pos_rfrc$density, group = "CXCL9+ FRC"),
  data.frame(density = cxcr3_pos_at_cxcl9_neg_rfrc$density, group = "CXCL9- FRC")
)

#### Plot the comparison between CXCL9+ and CXCL9- FRCs
density_df_fl$group <- factor(density_df_fl$group, levels = c("CXCL9+ FRC", "CXCL9- FRC"))
panel_j_lowerleft <- ggplot(density_df_fl, aes(x = group, y = density)) +
  geom_boxplot(aes(fill = group), outlier.shape = NA, width = 0.5, alpha = 0.7) +
  #geom_jitter(shape = 16, position = position_jitter(0.2), alpha = 0.4) +
  scale_fill_manual(values = c("#2AC200", "#f0f0f0")) + # Adjust the colors
  ylab("CXCR3+ T Density")+
  coord_flip()+
  stat_compare_means(method = "wilcox.test", label = "p.format", comp = list(c("CXCL9+ FRC", "CXCL9- FRC")), p.adjust.method = "fdr") +
  theme_bw() +
  theme(
    plot.title = element_blank(),
    plot.margin = margin(10, 10, 10, 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(color = "black"),
    legend.position = "none",
    aspect.ratio = 0.5,
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.x = element_text(size = 12, color = "black"),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = 12, color = "black")
  )

### DLBCL

cxcl9_pos_rfrc <- dlbcl_df %>% filter(cell_type_binned_cxcr3 == "CXCL9+ FRC")
cxcl9_neg_rfrc <- dlbcl_df %>% filter(cell_type_binned_cxcr3 == "CXCL9- FRC")
cxcr3_pos_t <- dlbcl_df %>% filter(cell_type_binned_cxcr3 == "CXCR3+ T")
#cxcr3_neg_t <- dlbcl_df %>% filter(cell_type_binned_cxcr3 == "CXCR3- T")

#### Convert coordinates into spatial point patterns (ppp format)
buffer <- 100  # Increase buffer slightly

x_min <- min(dlbcl_df$nucleus_centroid_x) - buffer
x_max <- max(dlbcl_df$nucleus_centroid_x) + buffer
y_min <- min(dlbcl_df$nucleus_centroid_y) - buffer
y_max <- max(dlbcl_df$nucleus_centroid_y) + buffer

window <- owin(xrange = c(x_min, x_max), yrange = c(y_min, y_max))
print(window)

#### Convert cell positions to spatial point patterns
cxcr3_pos_t_ppp <- ppp(cxcr3_pos_t$nucleus_centroid_x, cxcr3_pos_t$nucleus_centroid_y, window = window)
cxcl9_pos_rfrc_ppp <- ppp(cxcl9_pos_rfrc$nucleus_centroid_x, cxcl9_pos_rfrc$nucleus_centroid_y, window = window)
cxcl9_neg_rfrc_ppp <- ppp(cxcl9_neg_rfrc$nucleus_centroid_x, cxcl9_neg_rfrc$nucleus_centroid_y, window = window)

#### Perform Kernel Density Estimation (KDE) for CXCR3+ T cells
sigma_value <- 50  # Adjust as needed
cxcr3_pos_density <- density(cxcr3_pos_t_ppp, sigma = sigma_value)

#### Extract CXCR3+ T cell density at CXCL9+ and CXCL9- rFRC locations
cxcr3_pos_at_cxcl9_pos_rfrc <- as.data.frame(spatstat.geom::interp.im(cxcr3_pos_density, cxcl9_pos_rfrc_ppp))
cxcr3_pos_at_cxcl9_neg_rfrc <- as.data.frame(spatstat.geom::interp.im(cxcr3_pos_density, cxcl9_neg_rfrc_ppp))

colnames(cxcr3_pos_at_cxcl9_pos_rfrc) <- "density"
colnames(cxcr3_pos_at_cxcl9_neg_rfrc) <- "density"

#### Create a combined dataframe for both CXCL9+ FRC and CXCL9- FRC for comparison
density_df_dlbcl <- rbind(
  data.frame(density = cxcr3_pos_at_cxcl9_pos_rfrc$density, group = "CXCL9+ FRC"),
  data.frame(density = cxcr3_pos_at_cxcl9_neg_rfrc$density, group = "CXCL9- FRC")
)

#### Plot the comparison between CXCL9+ and CXCL9- FRCs
density_df_dlbcl$group <- factor(density_df_dlbcl$group, levels = c("CXCL9+ FRC", "CXCL9- FRC"))
panel_j_lowerright <- ggplot(density_df_dlbcl, aes(x = group, y = density)) +
  geom_boxplot(aes(fill = group), outlier.shape = NA, width = 0.5, alpha = 0.7) +
  #geom_jitter(shape = 16, position = position_jitter(0.2), alpha = 0.4) +
  scale_fill_manual(values = c("#2AC200", "#f0f0f0")) + # Adjust the colors
  ylab("CXCR3+ T Density")+
  coord_flip()+
  stat_compare_means(method = "wilcox.test", label = "p.format", comp = list(c("CXCL9+ FRC", "CXCL9- FRC")), p.adjust.method = "fdr") +
  theme_bw() +
  theme(
    plot.title = element_blank(),
    plot.margin = margin(10, 10, 10, 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(color = "black"),
    legend.position = "none",
    aspect.ratio = 0.5,
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 12, color = "black"),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = 12, color = "black")
  ) 
```


# Panels K, L, M
```{r}
CL_obj <- readRDS(".../CoCulture_DLBCL.rds") #accessible via Zenodo
DefaultAssay(CL_obj_sub) <- "RNA"

## Panel K
DimPlot(CL_obj, group.by = "group", reduction = "umap")+
  scale_color_manual(values = colors_cond)+
  NoLegend()

## Panel L
plot_df <- FetchData(CL_obj, vars = c("IFNG", "sample", "group", "cell_type"), slot = "data")
#plot_df$group <- factor(plot_df$group, levels = c("B+T", "LNSC+T", "LNSC+B+T"))
plot_df$cell_type <- factor(plot_df$cell_type, levels = c("iFRC", "B", "T"))

ggplot(plot_df, aes(x = cell_type, y = IFNG, fill = cell_type)) +
  geom_violin(
    scale = "width",
    adjust = 9, #smooth
    trim = TRUE
  ) +
  #scale_y_continuous(limits = c(0,4.5))+
  scale_fill_manual(values=colors)+
  ggtitle("IFNG") +
  ylab("Expression")+
  #facet_wrap(~group)+
  theme_classic() +
  theme(plot.title = element_text(size = 12, color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "black"), 
        aspect.ratio=2,
        axis.text.y = element_text(size = 12, color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15, color = "black", vjust = 1), 
        axis.text.x = element_text(size = 12, color = "black", angle = 0, hjust = 0.5, vjust = 0.5),
        legend.position = "none") +
  stat_compare_means(
    comparisons = list(c("B", "T")),
    method = "wilcox.test",
    p.adjust.method = "BH",
    label = "p.format"
  )

## Panel M

lnsc <- subset(CL_obj, subset = group %in% c("LNSC", "LNSC+B", "LNSC+T", "LNSC+B+T"))
lnsc <- subset(lnsc, cell_type == "iFRC")

get_gsea_genes <- function(msigdb_table,
                           gsea_term){
  gsea_genes <- msigdb_table %>% as_tibble() %>% mutate(pathway = gsub("\thttp:", "", V1), genes=V7)
  gsea_genes <- gsea_genes %>% filter(pathway == gsea_term) %>% dplyr::select(genes) %>% unlist() %>% strsplit2("\t") %>% .[,-1]
}
ifng <- get_gsea_genes(msigdb_table = gene_sets, gsea_term = "HALLMARK_INTERFERON_GAMMA_RESPONSE")

scores <- list()
scores[[1]] <- ifng
lnsc <- AddModuleScore(object = lnsc, features = scores, ctrl = 5, name = c("ifng"))
lnsc[['Scores']] <- CreateAssayObject(data = t(x = FetchData(object = lnsc, vars = c("ifng1"))))

lnsc@meta.data$group <- factor(x = lnsc@meta.data$group, levels = c("LNSC", "LNSC+B", "LNSC+T", "LNSC+B+T"))
comparisons = list(c("LNSC+T", "LNSC+B"), c("LNSC+B", "LNSC+B+T"))

VlnPlot(lnsc, features = "ifng1", group.by = "group", pt.size = 0, y.max = 0.35) +
      #stat_compare_means(comparisons = comparisons, method = "wilcox.test", label = "p.signif", ref.group = ".all.", hide.ns = FALSE) +
      geom_boxplot(inherit.aes = T, width=0.1, outlier.size = 0, alpha = 0.6, notch = FALSE) + 
      #geom_hline(yintercept = mean(lnsc$ifng1), linetype = 2) + # Add horizontal line at base mean
      geom_hline(yintercept = median(lnsc@meta.data$ifng1[lnsc@meta.data$group == "LNSC"]), linetype = 2) +
      ggtitle("LNSC ~ IFN response") + 
      scale_y_continuous(limits=c(-0.2, 0.3), breaks=c(-0.2,-0.1,0,0.1,0.2,0.3))+
      #coord_flip()+
      theme(plot.title = element_text(size = 12, color = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_rect(color = "black"), 
            aspect.ratio=1.5,
            axis.text.y = element_text(size = 12, color = "black"),
            axis.title.x = element_text(size = 12, color = "black"),
            #axis.title.y = element_text(size = 15, color = "black", vjust = 1), 
            axis.title.y = element_blank(), 
            axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1, vjust = 1)) + 
      labs(x = "", y = "Module Score\n") +
      stat_compare_means(
            method = "wilcox.test",
            p.adjust.method = "BH",
            label = "p.format",
            label.y = c(0.2,0.25),
            comparisons = comparisons
     ) +
      scale_fill_manual(values = colors_cond) +
      NoLegend()
```
