---
title: "Figure4"
output: html_document
date: "2024-02-29"
---
---
title: "Figure 4"
output: html_document
---


# Load packages
```{r setup, include=FALSE}
library(Seurat)
library(data.table)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(tidyverse)
library(ggalluvial)
```

# Load objects
For the following code please downlaod the following objects from Zenodo <insert link>:
- single cell LN stromal cell object: col 
- interaction potential object
```{r setup, include=FALSE}
scRNA <- readRDS("") #Available in zenodo
# Scriabin data
interaction_potential <- readRDS("path/to/interaction_potential.rds")
source("/path/to/Figure_config.R")
```

# Panels 
## A 
no need to show code 

## B
The interaction potential was calculated as stated in https://github.com/BlishLab/scriabin/blob/main/vignettes/single-dataset.Rmd and as data base Chemokines from the CytoSig data base were used. The calculation was done per patient and merged into one object interaction_potential.rds.

```{r Panel B, fig.height=8, fig.width=8, fig.cap="UMAP interaction potential"}
DimPlot(interaction_potential, group.by = "zones", cols=c("orange2", "#368F8A", "grey70", "#C0C26D"))+
  scale_x_reverse()+scale_y_reverse()+theme_void()
```

## C
```{r Panel C, fig.height=6, fig.width=12, fig.cap="UMAP feature plot color coded by the geometic mean expression of chemokines"}}
FeaturePlot(interaction_potential, features = c("CXCL13=CXCR5","CCL19=CCR7","CCL21=CCR7",
                                     "CXCL9=CXCR3","CXCL10=CXCR3","CXCL11=CXCR3"),
            split.by = "sender_group", cols = annotation_colors$features)
```

## D
```{r}
interaction_potential$sender_new_v5 <-interaction_potential$sender_new_v3
interaction_potential$sender_new_v5[interaction_potential$sender_Clusters_final %in% c("TRC", "PvC", "SMC")] <- "T-FRC"
interaction_potential$sender_new_v5[interaction_potential$sender_Clusters_final %in% c("MRC/IFR", "FDC")] <- "B-FRC"
interaction_potential$sender_new_v5[interaction_potential$sender_Clusters_final %in% c("med-FRC")] <- "med-FRC"
interaction_potential$sender_new_v5[interaction_potential$sender_Clusters_final %in% c("LAF")] <- "rFRC"

interaction_potential$sender_receiver_v5 <- paste0(interaction_potential$sender_new_v5, "_", interaction_potential$receiver_new_v3)

data <- interaction_potential@meta.data %>%
  dplyr::select(zones, seurat_clusters, sender_new_v5, receiver_new_v3, sender_group, sender_receiver_v5) %>%
  group_by(zones, sender_group, sender_new_v5, receiver_new_v3) %>% tally()

p1 <- as.data.frame(data) %>% dplyr::filter(zones=="B cell zone like ngh") %>% dplyr::filter(sender_group=="rLN") %>% 
  mutate(q50 = quantile(n, 0.5)) %>% dplyr::filter(n>unique(q50)) %>% 
  ggplot(aes(y = n, axis1 = sender_new_v5, axis2 = receiver_new_v3)) +
  geom_alluvium(aes(fill = sender_new_v5), width = 1/12)+
  geom_stratum(width = 1/12, fill = "grey70", color = "black") +
  #geom_label(stat = "stratum", aes(label = after_stat(stratum)))+
  scale_fill_manual(values = annotation_colors$Clusters_paper)+
  ggrepel::geom_text_repel(
    aes(label = ifelse(after_stat(x) == 1, as.character(after_stat(stratum)), NA)),
    stat = "stratum", size = 3, direction = "y", nudge_x = -5
  )  +  ggrepel::geom_text_repel(
    aes(label = ifelse(after_stat(x) == 2, as.character(after_stat(stratum)), NA)),
    stat = "stratum", size = 3, direction = "y", nudge_x = +5
  )+theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        #legend.position = "bottom",
        panel.background = element_rect(colour = "grey60", size=0.5))+
  xlab("")+
  ylab("")

p3 <- as.data.frame(data) %>% dplyr::filter(zones=="T cell zone like ngh") %>% #dplyr::filter(sender_group=="rLN") %>% 
    mutate(q50 = quantile(n, 0.50)) %>% dplyr::filter(n>unique(q50)) %>% 
  ggplot(aes(y = n, axis1 = sender_new_v5, axis2 = receiver_new_v3)) +
  geom_alluvium(aes(fill = sender_new_v5), width = 1/12)+
  geom_stratum(width = 1/12, fill = "grey70", color = "black") +
  #geom_label(stat = "stratum", aes(label = after_stat(stratum)))+
  scale_fill_manual(values = annotation_colors$Clusters_paper)+
  ggrepel::geom_text_repel(
    aes(label = ifelse(after_stat(x) == 1, as.character(after_stat(stratum)), NA)),
    stat = "stratum", size = 3, direction = "y", nudge_x = -5
  )  +  ggrepel::geom_text_repel(
    aes(label = ifelse(after_stat(x) == 2, as.character(after_stat(stratum)), NA)),
    stat = "stratum", size = 3, direction = "y", nudge_x = +5
  )+theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        #legend.position = "bottom",
        panel.background = element_rect(colour = "grey60", size=0.5))+
  xlab("")+
  ylab("")
```

## E
For this code the k6Data_rLN_percent_pairs.csv data set from Zenodo is needed
```{r}
k6Data_rLN <- read_csv("path/to/k6Data_rLN_percent_pairs.csv")

k6Data_rLN$summary_zone <- k6Data_rLN$id_nhood.x
k6Data_rLN$summary_zone[k6Data_rLN$id_nhood.x %in% c("BEC-rich T zone", "T zone")] <- "T zone"
k6Data_rLN$summary_zone[k6Data_rLN$id_nhood.x %in% c("B prol niche", "Follicle cluster 1")] <- "B zone"

k6Data_rLN_summary <- k6Data_rLN %>% group_by(summary_zone, pair) %>% summarise(n_pair_summary= sum(n_pair), 
                                                          n_pair_total= sum(n_total))
data <- k6Data_rLN %>%
  filter(pair %in% c("FDC_B", "T-FH_B", "MSC_B", "MSC_T-H", "MSC_T-TOX", "MSC_T-REG", "MSC_Myeloid")) %>%
  filter(!id_nhood.x %in% c("Plasma cells","Lymphatic sinus")) %>%
  dplyr::select(id_nhood.x, pair, pct_codex) %>%
  pivot_wider(values_from = "pct_codex", names_from = pair) %>% column_to_rownames("id_nhood.x") %>%
  scale() %>%
  t()

k6Data_rLN_summary <- k6Data_rLN %>% group_by(summary_zone, pair) %>% summarise(n_pair_summary= sum(n_pair), 
                                                                                n_total_summary = sum(n_total))
k6Data_rLN_summary$fraction <- k6Data_rLN_summary$n_pair_summary/k6Data_rLN_summary$n_total_summary
data <- k6Data_rLN_summary %>% 
    filter(pair %in% c("FDC_B", "T-FH_B", "MSC_B", "MSC_T-H", "MSC_T-TOX", "MSC_T-REG", "MSC_Myeloid")) %>%
    filter(summary_zone %in% c("T zone","B zone")) %>% 
    dplyr::select(summary_zone, pair, fraction) %>%  
    pivot_wider(values_from = fraction, names_from = pair) %>% column_to_rownames("summary_zone") %>% 
    #scale() %>% 
    t()

data_scriabin <- interaction_potential@meta.data %>% 
  group_by(sender_group, zones, sender_receiver_v3) %>% tally() %>% 
  ungroup() %>% group_by(zones) %>%  mutate(Frac = n/sum(n)) %>% 
  filter(sender_group=="rLN") %>% 
  filter(sender_receiver_v3 %in% c("FDC_B", "T-FH_B", "MSC_B", "MSC_T-H", "MSC_T-TOX", "MSC_T-REG", "MSC_Myeloid")) %>%
  filter(!zones %in% c("other")) %>% 
  dplyr::select(zones, sender_receiver_v3, Frac) %>%  
  pivot_wider(values_from = "Frac", names_from = sender_receiver_v3) %>% column_to_rownames("zones")


data_joined <- t(data_scriabin) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  left_join(rownames_to_column(as.data.frame(data))) %>% 
  column_to_rownames()

#data_joined$`B cell zone like ngh`[5:6] <- 0.000001

data_joined <- data_joined[,-2]

pheatmap::pheatmap(data_joined, 
          color =colorspace::diverging_hcl(n = 40, palette = "Blue-Red3"),
          #rev(colorspace::diverging_hcl(n = 40, h = c(-324, 193), c = 100, l = c(60, 100), power = 1)), 
          scale = "row", cluster_cols = T, cluster_rows = T, na_col = "grey80")+coord_flip()
  

```


## F
```{r}
interaction_potential$UMAP_1 <- interaction_potential@reductions$umap@cell.embeddings[,1]
interaction_potential$UMAP_2 <- interaction_potential@reductions$umap@cell.embeddings[,2]
interaction_potential@meta.data %>%
  arrange(factor(sender_group, levels = c("FL","DLBCL","rLN"))) %>% ggplot() + 
  scattermore::geom_scattermore(mapping = aes_string(x = "UMAP_1", 
                                                     y = "UMAP_2", color = "sender_group"))+
  scale_color_manual(values = c("#8D60B5", "grey85","grey85"))+
  scale_x_reverse()+
  scale_y_reverse()+
  theme_void()

interaction_potential@meta.data %>%
  arrange(factor(sender_group, levels = c("rLN","DLBCL","FL"))) %>% ggplot() + 
  scattermore::geom_scattermore(mapping = aes_string(x = "UMAP_1", 
                                                     y = "UMAP_2", color = "sender_group"))+
  scale_color_manual(values = c("grey85", "#00A1E0","grey80"))+
  scale_x_reverse()+
  scale_y_reverse()+
  theme_void() ###color changed bechause it is to light

interaction_potential@meta.data %>%
  arrange(factor(sender_group, levels = c("FL","rLN","DLBCL"))) %>% ggplot() + 
  scattermore::geom_scattermore(mapping = aes_string(x = "UMAP_1", 
                                                     y = "UMAP_2", color = "sender_group"))+
  scale_color_manual(values = c("grey80", "grey80","#67B86E"))+
  scale_x_reverse()+
  scale_y_reverse()+
  theme_void()
```

## G
### left
```{r}
interaction_potential@meta.data %>% 
  group_by(zones,sender_group) %>% tally() %>% 
  #filter(!zones=="other") %>% 
  ggplot(aes(x=sender_group,
             y=n, fill=factor(zones, levels = rev(c("T cell zone like ngh", "B cell zone like ngh", "Inflammatory ngh", "other")))))+
  geom_bar(position="fill",stat="identity")+
  scale_fill_manual(values = rev(c("#5e65be",  "#74c8c3", "#d8d97a", "grey70")), name="LN zones")+

  theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(colour = "grey60", size=0.5))+
  xlab("")+
  ylab("Scaled fractions per zone")

```

### right 
```{r}
interaction_potential$sender_new_v4 <- ifelse(interaction_potential$sender_Clusters_final %in% c("FDC", "med-FRC", "MRC/IFR", "PvC", "SMC", "TRC"), 
                                   "FRC", merged_obj$sender_Clusters_final )
interaction_potential$reciver_new_v4 <- ifelse(interaction_potential$receiver_Clusters_final %in% c("FDC", "med-FRC", "MRC/IFR", "PvC", "SMC", "TRC"), 
                                   "FRC", interaction_potential$receiver_Clusters_final )
interaction_potential$sender_reciver_v4 <- paste0(merged_obj$sender_new_v4, "_", merged_obj$reciver_new_v4  )

freq <- interaction_potential@meta.data %>% group_by(zones, sender_patient,sender_group) %>% 
  count(sender_reciver_v4) %>% ungroup() %>% group_by(sender_patient, sender_group) %>%     
  mutate(count=sum(n)) %>% ungroup() %>% filter(zones %in% c("Inflammatory ngh")) %>%
  dplyr::select(!zones) %>% group_by(sender_patient, sender_group, sender_reciver_v4, count) %>% summarise(sum_n=sum(n)) %>% 
  mutate(rel_count=sum_n/count) %>% ungroup() %>% group_by(sender_reciver_v4) %>% mutate(sum=sum(rel_count))
cutoff <- quantile(freq$sum, 8/10)
freq <- freq %>% filter(sum>0.30)

ggplot(freq, aes(x=reorder(sender_reciver_v4, sum) , y=rel_count, fill=sender_group)) + 
  geom_bar(stat='identity') + theme_classic() + scale_fill_manual(values = annotation_colors$Type_binned) +
  xlab("")+coord_flip()
```


#H
```{r Panel H, fig.height=4, fig.width=5, fig.cap="Violin plot showing the difusness score"}
interaction_potential_sampled <- readRDS("/omics/odcf/analysis/OE0285_projects/human-bone-marrow/Lymphoma/bd_rhapsody/2021_May/Analysis/LIANA/merged_obj_sub_downsampled_rowmean_pca_0.3.rds")

ggplot(interaction_potential_sampled@meta.data, aes(y= rowmean_pca, x = sender_group, fill=sender_group))+
  geom_violin(scale = "width")+#, pt.size = 0)+
  geom_boxplot(width=0.2, outlier.size = 0)+
  scale_fill_manual(values = annotation_colors$Type_binned, name="disease")+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    panel.background = element_rect(colour = "grey60", size=0.5))+
  scale_y_log10()+
  xlab("")+
  ylab("Diffusness score")+
  ggtitle("Diffusness score")+NoLegend()

interaction_potential_sampled@meta.data %>% rstatix::t_test(rowmean_pca~sender_group)
```


# I
```{r Panel I, fig.height=4, fig.width=5, fig.cap="Violin plot showing the R-L of homeostatic chemokines"}
merged_obj <- AddModuleScore(merged_obj, features = list(c("CXCL13=CXCR5", "CXCL12=CXCR4", "CCL19=CCR7", "CCL21=CCR7"),
                                                                 c("CXCL9=CXCR3" ,"CXCL10=CXCR3" , "CXCL11=CXCR3")), 
                                 name = c("homeostatic", "ifn"), nbin = 2, ctrl = 20)

ggplot(interaction_potential@meta.data, aes(y=homeostatic1 , x = sender_group, fill=sender_group))+
  geom_violin()+#, pt.size = 0)+
  geom_boxplot(width=0.2, outlier.size = 0)+
  scale_fill_manual(values = annotation_colors$Type_binned, name="disease")+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    #legend.position = "bottom",
    panel.background = element_rect(colour = "grey60", size=0.5))+
  xlab("")+
  ylab("Organizing score")+
  ggtitle("Organizing score")+NoLegend()

```

#J
```{r  Panel J, fig.height=4, fig.width=5, fig.cap="Violin plot feature plot showing the R-L of homeostatic chemokines"}
FeaturePlot(interaction_potential, features = c("ifn2"), cols = annotation_colors$features)

ggplot(interaction_potential@meta.data, aes(y= inf2, x = sender_group, fill=sender_group))+
  geom_violin(scale = "width")+#, pt.size = 0)+
  geom_boxplot(width=0.2, outlier.size = 0)+
  scale_fill_manual(values = annotation_colors$Type_binned, name="disease")+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    #legend.position = "bottom",
    panel.background = element_rect(colour = "grey60", size=0.5))+
  xlab("")+
  ylab("Inflammotory score")+
  ggtitle("Inflammotory score")+NoLegend()
```